### 2022-01-18 manuscript.tex
initial commit pre new figs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx} % \usepackage[draft]{graphicx}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\begin{document}

\title[]{Doublet splitting and linear amplitude characteristics of fusion alpha-particle driven ion cyclotron emission}

\author{J. W. S. Cook}

\address{UKAEA-CCFE, Culham Science Centre, Abingdon, OX14 3DB, UK}
\ead{james.cook@ukaea.uk}
\vspace{10pt}
\begin{indented}
\item[]Jan 2022
\end{indented}

\begin{abstract}
Ion cyclotron emission (ICE) originating from confined populations of fast ions in toroidal fusion plasmas is an important non-invasive, passive diagnostic for current and next generation devices. The ability to model the ICE signals accurately is an essential step towards inferring the characteristics of confined energetic alpha-particles or fast NBI ions and background plasma. In this paper, the linear growth rates of the magnetoacoustic cyclotron instability, which is the leading explanation for ICE, are calculated in high resolution 2D $(k_\parallel, k_\perp)$ space for parameters corresponding to JET Pulse No. 26148. These calculations shed further light on the origin of doublet-splitting of the cyclotron harmonics as well as the apparent need to invoke nonlinear interactions with 1D3V particle-in-cell studies to account for lower harmonics with substantial amplitudes: the inclusion of signals of ICE from a well resolved range of propagation angles accounts for these effects.
\end{abstract}

%
% Uncomment for keywords
%\vspace{2pc}
%\noindent{\it Keywords}: XXXXXX, YYYYYYYY, ZZZZZZZZZ
%
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
%
% Uncomment if a separate title page is required
%\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

%%begin novalidate

\section{Introduction}

Observations of ion cyclotron emission (ICE) have been reported from magnetically confined plasmas first in TFTR\cite{Cauffman1995,Dendy1995} and JET\cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks PDX\cite{Heidbrink1994}, JT-60-U\cite{Ichimura2008}, DIII-D\cite{Heidbrink2011}, ASDEX\cite{DInca2014}, ASDEX-Upgrade\cite{Ochoukov2018}, EAST\cite{Liu2020}, and KSTAR\cite{Chapman2017,Chapman2018}. Stellarators\cite{Saito2013} and FRCs\cite{Nicks2020} have also exhibited this phenomenon. Energetic ion populations that give rise to ICE have their origins in fusion reactions in DT plasmas\cite{Cottrell2000, Cauffman1995}, ICRH\cite{Cottrell2000}, and neutral beam injection\cite{Ochoukov2018}. ICE arising from fusion $\alpha$-particles is a proposed route to further the understanding of burning plasmas in ITER\cite{Dendy2015,McClements2015}.

The foremost explanation of the mechanism responsible for ICE is the magnetoacoustic cyclotron instability (MCI)\cite{Belikov1976}, which arises from the free energy present in energetic ions with inverted distributions in velocity space. Theoretical understanding of ICE is based in analytical solutions to the linearised Maxwell-Vlasov system of equations\cite{Stix} where the energetic ion species was modelled as ring-beam for which $f(v_\parallel, v_\perp) \propto \exp(-(v_\parallel-u_\parallel)^2/ v_{th}^2)\delta(v_\perp - u_\bot)$ \cite{Dendy1994, McClements1996}. Here $u_\parallel$, $u_\perp$ and $v_{th}$ are the beam velocity (the signed drift speed along the magnetic field), ring speed (the speed perpendicular to the magnetic field) and thermal velocity, respectively. Subsequent studies augmented this understanding via nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one dimension and 3 vector components for velocities and electromagnetic field (i.e. 1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These computational approaches also operate under the simplification that the plasma is both homogeneous and infinite albeit approximately via periodic boundary conditions. Linearised fluid approaches that incorporate toroidal geometry and inhomogeneous plasma and magnetic fields have shed light of the eigenmode structure of compressional Alfv{\'e}n eigenmodes, which like ICE are also fast Alfv{\'e}n modes driven by energetic ions, in spherical tokamaks\cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio tokamaks \cite{Kolesnichenko2000, Fulop2000, Heidbrink2006}.

The effectiveness of analytical theory to reproduce the observed cyclotron harmonic peaks in the fluctuations of the magnetic field was first explained somewhat by 1D PIC simulations where the onset of nonlinearity appeared to imprint into the signal the amplitudes of each of the harmonics as they are at the end of the linear phase; each mode grows independently until growth is terminated for all modes when the fastest mode enters its nonlinear phase. Analysis shows that only higher cyclotron harmonics are destabilised\cite{McClements2015} to the MCI, however observations clearly indicate power at theoretically stable lower harmonics\cite{Cottrell1988}. In a further step towards accurate recreation of experimental observations, nonlinear phases of hybrid\cite{Carbajal2014} and fully kinetic\cite{Cook2013,Chapman2017} PIC simulations went on to show that wave-wave coupling was responsible for raising the amplitude of lower cyclotron harmonics. Most recently, hybrid 2D simulations have begun to shed light on the 2D3V dynamics of ICE\cite{Carbajal2021}.

To better enable ICE as a diagnostic for confined fast ions, it is necessary to further develop our understanding with a range of tools. In this work, I present the 2D $(k_\parallel, k_\perp)$ growth rates of the MCI, under the conditions of the JET PTE\cite{Cottrell1993} namely Pulse No. 26148 are presented. These growth rates are obtrained from a dispersion relation solver, PlasmaDispersionRelations.jl, written in the Julia programming language\cite{Bezanson2017}. These results narrow the gap between analytical theory and expensive HPC simulations and can be run readily on desktop computers in a matter of minutes, further enabling research into this topic.

In section 2, the computational method is described for calculating the linear growth rates, which are presented for energetic ring-beam alpha-particles with non-zero perpendicular thermal spread in section 3. Results are summarised in section 4.

\section{Computational method}

Solutions to the linearised Maxwell-Vlasov system of equations are found in $(k_\parallel, k_\bot)$ space for a ring-beam distribution of energetic minority alpha-particles and Maxwellian electrons and background thermal majority deuteron ions. Calculations are made for alpha-particles with ring-beams of two pitch values: $u_\parallel=-0.64 v_0 \simeq -1.085 V_A$, which is associated with the ion-cyclotron emission observed on JET Pulse No. 26148, see Figs. 3 and 15 and associated text in Ref. \cite{Cottrell1993}; and $u_\parallel=0 v_0$ representing a ring distribution with no parallel drift, see Fig. 1(a) of Ref \cite{Cook2013}. Here $v_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is the speed of an alpha-particle with an energy of 3.6 MeV. The complex frequency solutions, $\omega$, are found corresponding to the fast Alfv{\'e}n branch on a grid of $(k_\parallel, k_\bot)$.

The plasma parameters approximate those found in the region indicated in Fig. 3 of Ref. \cite{Cottrell1993}, which are associated with the continuous trace signal of Fig. 2 (ibid). The electrons and deuterons, both modelled as Maxwellians, have temperatures of 1 keV and the electrons have a number density of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$ $\mathrm{T}$.
The ring-beam distribution function, Eq. \ref{eq:alphaf}, is a separable function of $v_\parallel$, i.e. the beam given by Eq. 1, and $v_\perp$, i.e. the ring given by Eq. 2. The component parallel (perpendicular) to the magnetic field is a drifting Maxwellian with thermal speed $v_{\parallel, th}$ ($v_{\perp, th}$) and drift velocity $u_\parallel$ ($u_\perp$). The thermal spread of the alpha particles' distribution function $v_{th}=v_{\parallel,th}=v_{\perp,th}=v_0/100$. The concentration ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.

\begin{align}
f_{\alpha,\parallel}\left(v_\parallel\right) &= \frac{\exp\left(-\frac{(v_\parallel - u_\parallel)^2}{v_{\parallel,th}^2}\right)}{v_{\parallel,th} \sqrt{2 \pi}}\\
f_{\alpha,\perp}\left(v_\perp\right) &= \frac{1}{2\pi}\frac{\exp\left(-\frac{\left(v_\bot - u_\bot\right)^2}{v_{\bot,th}^2}\right)}{\left(\frac{\sqrt \pi}{2} v_{\perp, th} u_\bot (1 - \mathrm{erf}(-\frac{u_\bot}{v_{\perp,th}})) + \exp(-u_\bot^2 / v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}\\
f_\alpha\left(v_\parallel, v_\perp\right) &= f_{\alpha,\parallel}\left(v_\parallel\right)f_{\alpha,\perp}\left(v_\perp\right)
    \label{eq:alphaf}
\end{align}

Solutions to the linearised Maxwell-Vlasov system of equations are given by the dispersion relation, Eq A.1.
The hot dielectric tensor\cite{Stix} given by Eq. A.2 is calculated for each species.

The integrals over perpendicular velocity required for calculating the dielectric tensor for Maxwellian distribution functions are given in terms of modified Bessel functions of the first kind for both deuterons and ions i.e. when $u_\perp =0$. The same perpendicular integrals over the alpha-particles' ring distribution is calculated numerically via Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK} library. An absolute and relative tolerance of $0$ and $10^{-8}$, respectively, on the value of the integrals is used.

Since all the parallel distribution functions for all species are Maxwellians, the parallel integrals can be calculated from the well known plasma dispersion function\cite{Fried1961}, Eqs. B.1-B.3.

The summation over Bessel function indices continues until the relative change of the L2 norm between the current and previous iteration is less than a relative tolerance of $10^{-8}$. Root finding of Eq. A.1 is performed by an amalgamation of the Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method, whereby if the Nelder-Mead simplex does not contain a root, as calculated via the winding method, then the optimisation procedure follows the Nelder-Mead method, otherwise if a root is present with the simplex then the simplex bifurcates and the child simplex that contains the root is used for the next iteration. The optimisation iteration procedure finishes when the real extent of the simplex is less than $10^{-4}$ of the alpha-particle cyclotron frequency $\Omega_i$ (the only ion species considered are deuterons and alpha-particles, which have identical cyclotron frequencies) and the imaginary extent is less than $10^{-5}\,\Omega_i$. This approach is found to be adequately robust and faster than the Nelder-Mead method or winding number approach used in isolation of one another. Each 512 by 512 scan in $(k_\parallel, k_\perp)$ takes between 5 and 6 minutes on a laptop.

\section{Linear growth rates}

The growth rates for the zero-pitch case ($u_\parallel = 0 v_0$ in Eq. 1) are plotted in Fig. \ref{fig:2D_F12_zeropitch} against the associated real frequency (damped modes are not shown). In this figure shading indicates the propagation angle, where $90^\circ$ ($0^\circ$) is exactly perpendicular (parallel) to the magnetic field. Fig. \ref{fig:2D_imag_zeropitch} plots shows the same data as Fig. \ref{fig:2D_F12_zeropitch} except shading indicates the growth rates in $(k_\parallel, k_\perp)$ space. Overlaid contours indicate the real frequency in units of ion cyclotron harmonics. Only narrow regions around the cyclotron harmonics are unstable. Overlaid straight traces indicate propagation angle. No Doppler shift is present due to the lack of parallel drift in the alpha-particle ring-beam. The growth rate increases approximately linearly with cyclotron harmonic number, which is consistent with the literature\cite{Dendy1994}. Together these figures serves to introduce a 2D view (growth rates in $(k_\parallel, k_\perp)$ space) of traditionally 1D data (growth rates against frequency). This will become important when a non-zero pitch is taken into account next.

\begin{figure}[ht!]
    \centering
\includegraphics[scale=0.8]{ICE2D_F12_zeropitch.pdf}
    \caption{(colour online) The imaginary component of the frequencies of the unstable fast Alfv{\'e}n wave solutions to the linear Maxwell-Vlasov system of equations, Eq. A.1, for thermal electrons, deuterons and a ring distribution of energetic alpha-particles i.e. when there is zero parallel drift, $u_\parallel=0$ in Eq. 1. Consecutive cyclotron harmonics are increasingly unstable and are narrow due to the lack of Doppler shift associated with parallel drifts. Shading indicates the propagation angle of the complex frequency solution. These data are also shown in Fig. \ref{fig:2D_imag_zeropitch} in $(k_\parallel,k_\perp)$ space. \ref{fig:2D_real_zeropitch}.}
    \label{fig:2D_F12_zeropitch}
\end{figure}

\begin{figure}[ht!]
    \centering
\includegraphics[scale=0.8]{ICE2D_imag_zeropitch.pdf}
    \caption{(colour online) Shading indicates the growth rate of fast Alfv{\'e}n waves excited by a ring of fusion born alpha-particles, as a function of parallel wavenumber on the ordinate and perpendicular wavenumber on the abscissa. No solution is found where shading is absent. The alpha-particle distribution function is given by Eq. \ref{eq:alphaf} with $u_{\parallel} = 0$. Consecutive straight traces indicate the propagation angle with respect to the magnetic field: the horizontal line is exactly perpendicular; solid traces fan out at intervals of five degrees; and dashed traces indicate intervals of one degree. Curved grey traces indicate the location in $(k_\parallel, k_\perp)$ of the cyclotron harmonics (i.e. the real counterpart to the imaginary frequency), which are annotated along the left hand side and upper edge of the figure. Frequencies are in units of the alpha-particle cyclotron frequency, and wavenumbers are in alpha-particle cyclotron frequencies per Alfv{\'e}n speed. Note the up-down symmetry due to the lack of parallel drift in the energetic alpha particles.}
    \label{fig:2D_imag_zeropitch}
\end{figure}

Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch case ($u_\parallel = -0.64 v_0$) in the same manner as Fig. \ref{fig:2D_F12_zeropitch}. There are several points worth noting. First, the growth rate is no longer linear with cyclotron harmonic number. Second, harmonics are broader due to the non-zero parallel drift that gives rise to strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting of the $8^{th}$ harmonic, for example, is observed as it is in Fig. 2 of Cottrell 1993 and as is indicated by Fig. 3 of Dendy et al 1994 (also Fig. 9 Dendy et al 1995), which shows slices at constant propagation angle that when all superimposed point towards doublet splitting, albeit under-resolved in wavenumber space. The following describes how these three features are connected. The doublet splitting arises due to the dip in growth rate where $k_\parallel$ goes through zero combined with the change in sign of the Doppler shift of the real frequency as $k_\parallel$ changes sign. The effect is exaggerated for higher harmonics where the growth rate diminishes to zero for perpendicular propagation; see e.g. the $10^{th}$ harmonic in Fig. \ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A, k_\bot \simeq 11\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron frequency and $V_A$ is the Alfv{\'e}n speed. Other possible mechanisms for doublet splitting could be grad-B or curvature drifts\cite{Cottrell1993}. Qualitatively, Fig. \ref{fig:2D_F12_physicalpitch} resembles Fig. 2 of Cottrell 1993 in that the signal below $\sim 7$ cyclotron harmonics is lower and flatter with well defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change larger and the harmonics are less easily identifiable. These features can only be recreated with a 2D calculation for these distributions functions. Note that Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes more clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is characterised by narrow regions of instability near Doppler shifted cyclotron harmonics that are stabilised when $k_\parallel=0$ but unstable at larger negative parallel wavenumbers. The higher frequency behaviour in the region $\omega\geq 7\Omega_i$ is dominated by broader regions of instability in frequency and wavenumber near to Doppler shifted cyclotron harmonics, which have a complex structure in wavenumber space. Modes propagating exactly perpendicular become stable in the region $k_\perp > 13 \Omega_i/V_A$ whilst obliquely propagating modes remain unstable; see Fig. 9 of Dendy et al 1995. The Doppler shift of the growing modes can reach $\approx \Omega_i$ for large propagation angles at lower frequencies and also for moderately oblique angles at higher frequencies; see, for example, the mode associated with the $15^{th}$ cyclotron harmonic at, $(k_\parallel \simeq -\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e. a propagation angle of $93^\circ$). Furthermore, the breadth of the harmonics can be attributed to the contributions from multiple propagation angles even though the lower harmonics are stable for forward travelling and perpendicular propagation angles.

\begin{figure}[ht!]
    \centering
\includegraphics[scale=0.8]{ICE2D_F12_physicalpitch.pdf}
    \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 v_0$. See also Fig. \ref{fig:2D_imag_physicalpitch} plotted against the real part, shown in \ref{fig:2D_real_zeropitch}. Shading indicates the wave's propagation angle. Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks. The 8th cyclotron harmonic is split into a doublet, or has two peaks because both the growth rate and real frequency vary with propagation angle. Compare with Fig. 2 of Cottrel et al 1993.}
    \label{fig:2D_F12_physicalpitch}
\end{figure}

\begin{figure}[ht!]
    \centering
\includegraphics[scale=0.8]{ICE2D_imag_physicalpitch.pdf}
    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch value resulting in $u_{\parallel} = -0.64 v_0$. The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch of the opposite sign, $u_{\parallel} = 0.64 v_0$, is identical expect flipped about $k_\parallel=0$.}
    \label{fig:2D_imag_physicalpitch}
\end{figure}

\begin{figure}[ht!]
    \centering
\includegraphics[scale=0.8]{ICE2D_real_div_guess_zeropitch.pdf}
    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except that the shading indicates the real part of the smoothed solution normalised by the high frequency expression for the fast Alfv{\'e}n wave set out by Eq. 25 of Ref. \cite{Dendy1994}, which is also Eq. 11 in Ref. \cite{McClements1996}. The differences in real frequency for the two sets of calculations, physical pitch of $-0.64$ and zero pitch, are negligible.}
    \label{fig:2D_real_zeropitch}
\end{figure}

\section{Summary}

This paper presents the linear Maxwell-Vlasov dispersion relation of a plasma consisting of thermal electrons, thermal deuteron fuel ions and a minority energetic ring-beam of fusion born alpha-particles. These high resolution $(k_\parallel, k_\perp)$ calculations revisit ICE observations from JET Pulse No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features: the step change in signal strength from $\omega < 6$ to $\omega > 6$; the doublet splitting of peaks at $\omega \approx 10\Omega_i$ on Fig. \ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\parallel, k_\perp)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron frequency, which makes it difficult to determine which parts of the signal come from which cyclotron harmonic numbers. Further effort is required to account for the extra doublets visible in the continuous trace associated with JET Pulse No. 26148 in Fig. 2 of Ref. \cite{Cottrell1993}.

These calculations capture the finite perpendicular thermal spread of the alpha particle distribution function, the integral of which has no known analytically tractable form. The computational capabilities of the software, PlasmaDispersionRelations.jl, used to generate the results further enable ICE as a diagnostic for fusion plasmas since the code is capable of: calculating the dispersion relation of plasmas consisting of an arbitrary number of species; and of arbitrary distribution functions, although those with parallel Maxwellians, like those presented here, are faster. It has been shown that it is possible to recreate signals containing lower harmonics of substantial amplitudes by invoking 2D linear physics. The presence of ICE signals at lower harmonics in PIC simulations has been attributed to nonlinear interactions between pairs of more powerful higher harmonics\cite{Carbajal2014}. This work suggests that observations of lower ICE harmonics may arise either due to the effects of 2D linear physics or to nonlinear 1D physics\cite{Chapman2018}, or a combination. An important feature missing from these calculations is the effect of plasma geometry, which impacts the ICE signals according to frequency and propagation angle. Cottrell et al. \cite{Cottrell1993} does discuss the implications of propagation angles on ICE signals and indicate that lower harmonics must propagate increasingly close to exactly perpendicular. These effects remain the subject of further work to incorporate them in linear and nonlinear studies of ICE.

This work showcases a new high $(\omega,k)$ resolution tool for studying ICE and captures the fine detail of ICE signals, which is challenging to capture with nonlinear PIC code simulations since they suffer from low resolution in frequency and wavenumber space due to the phenomenon's transience and computational expense, respectively. As such, this work represents a complementary tool that builds upon purely analytical work and nonlinear PIC studies, which in turn enhances the feasibility of exploiting ICE as a passive diagnostic of confined and leaving ions in magnetised fusion plasmas.

\section{Acknowledgements}

It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou useful discussions.

\appendix
\section{}

The plasma dispersion relation\cite{Stix} is given by the $(\omega,\vec{k})$ solutions to A.1, given the contributions of the dielectric tensor $\vec{\vec{\epsilon}}$ from all species.

\begin{align}
    0 &= ||\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}||,\\
    \vec{\vec{\epsilon}} &= \vec{\vec{1}} + \sum_s \frac{\Pi_s^2}{\omega}\sum_{n=-\infty}^{\infty}
    \int_0^{\infty}
    \int_{-\infty}^{\infty}\frac{\vec{\vec{S}}_{s,n} 2\pi v_\perp dv_\perp dv_\parallel}{\omega - k_\parallel v_\parallel - n \Omega_s}
\end{align}

where

\begin{align}
\vec{\vec{S}}_{s,n} &=
\begin{bmatrix}
\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n J_n^2}{z_s} v_\perp W_s\\
\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp W_s \\
\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2 v_\parallel W_s \\
\end{bmatrix},\\
U_s &= \frac{\partial f_s}{\partial v_\perp} + \frac{k_\parallel}{\omega}  \left(v_\perp \frac{\partial f_s}{\partial v_\parallel} - v_\parallel \frac{\partial f_s}{\partial v_\perp}\right), \\
W_s &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp},\\
\end{align}

$J_n$ is the $n$th Bessel function of the first kind and takes the argument $z_s=\frac{k_\perp v_\perp}{\Omega_s}$ where $J_n'=\frac{\partial J_n}{\partial z}$.

\section{}

The plasma dispersion function\cite{Fried1961}

\begin{equation}
\it{Z}_n(z)=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}\frac{x^n e^{-x^2}}{x-z}dx,
\end{equation}

can be calculated via the recurrence relation\cite{Sampoorna2007}

\begin{equation}
{\it Z}_n(z) = z {\it Z}_{n-1}(z) + \begin{cases}
0 & \text{if } n \in \{2k : k \in \mathbb{Z}\}
\\
1 & \text{if } n = 1
\\
\frac{1}{2} & \text{if } n = 3
\\
2^{\frac{1-n}{2}} \Pi_{i=1}^{\frac{n-3}{2}} \left( 2i + 1 \right) & \text{otherwise},
\end{cases}
\end{equation}

where $n \geq 0$ and the cases $n=1$ and $n=3$ are stated explicitly to avoid confusion. The zeroth term is defined as

\begin{equation}
 {\it Z}_0(z) = i \sqrt{\pi} \mathrm{erfcx} \left(-i z \right)
\end{equation}

where $\mathrm{erfcx}$ is the scaled complementary complex error function.

\newpage

% \clearpage

%%end novalidate
\begin{thebibliography}{9}

\bibitem{Cauffman1995}
Cauffman S, Majeski R, McClements K G, and Dendy R O 1995 \textit{Nucl. Fusion} \textbf{35} 1597
% , "Alfvenic behaviour of alpha particle driven ion cyclotron emission in TFTR,"

\bibitem{Dendy1995}
Dendy R O, McClements K G, Lashmore-Davies C, Cottrell G A, Majeski R, and Cauffman S 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
%  , “Ion cyclotron emission due to collective instability of fusion prod-ucts and beam ions in TFTR and JET,”

\bibitem{Cottrell1988}
Cottrell G A and Dendy R O 1988 \textit{Phys. Rev. Lett.} \textbf{60} 33-6

\bibitem{Cottrell1993} 
Cottrell G A, Bhatnagar V, Costa O D, Dendy R O, Jacquinot J, McClements K G, McCune D, Nave M, Smeulders P, and Start D 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
% 
% "Ion cyclotron emission measurements during JET deuterium-tritium experiments,"

\bibitem{Heidbrink1994}
Heidbrink W and Sadler G 1994 \textit{Nucl. Fusion} \textbf{34} 535–615
% "The behaviour of fast ions in tokamak experiments,"

\bibitem{Ichimura2008}
Ichimura M, Higaki H, Kakimoto S, Yamaguchi Y, Nemoto K, Katano M, Ishikawa M, Moriyama S, and Suzuki T 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
% , "Observation of spontaneously excited waves in the ion cyclotron frequency range on JT-60u,"

\bibitem{Heidbrink2011}
Heidbrink W W et al 2011 \textit{Plasma Phys. Control. Fusion} {\textbf 53} 085028
% "Characterization of off-axis fishbones",

\bibitem{DInca2014}
D’Inca R 2014 \textit{Ion Cyclotron Emission on ASDEX Upgrade PhD Thesis} Ludwig-Maximilians-Universitat

\bibitem{Ochoukov2018}
Ochoukov R et al 2018 \textit{Rev. Sci. Instrum.} 2018 \textbf{89}, 10J101
% , "Observations of core ion cyclotron emission on ASDEX Upgrade tokamak"
% , Bobkov V, Chapman B, Dendy RO and Dunne,M  and Faugel, H  and García-Muñoz,M  and Geiger,B  and Hennequin,P  and McClements,K G  and Moseev,D  and Nielsen,S  and Rasmussen,J  and Schneider,P  and Weiland,M  and Noterdaeme,J-M

\bibitem{Liu2020}
Liu L et al 2020  \textit{Nucl. Fusion} \textbf{60} 044002
% "Ion cyclotron emission driven by deuterium neu- tral beam injection and core fusion reaction ions in EAST,"
% Zhang X, Zhu Y, Qin C, Y Zhao, S Yuan, Y Mao, M Li, K Zhang, J Cheng, L Ai, and Y Cheng

\bibitem{Chapman2017}
Chapman B, Dendy R O, McClements K, Chapman S C, Yun G, Thatipamula S, and Kim M 2017 \textit{Nucl. Fusion} \textbf{57} 124004
% , "Sub-microsecond temporal evolution of edge density during edge localized modes in KSTAR tokamak plasmas inferred from ion cyclotron emission,"

\bibitem{Chapman2018}
Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027

\bibitem{Saito2013}
Saito K et al 2013 \textit{Plasma Sci. Technol.} \textbf{15} 209–212
% "Measurement of ion cyclotron emissions by using high-frequency magnetic probes in the LHD," 
%R Kumazawa, T Seki, H Kasahara, G Nomura, F Shimpo, H Igami, M Isobe, K Ogawa, K Toi, M Osakabe, M Nishiura, T Watanabe, S Yamamoto, M Ichimura, and T Mutoh

\bibitem{Nicks2020}
Nicks B S, Magee R, Necas A and Tajima T 2021 \textit{Nucl. Fusion} \textbf{61} 016004
% "Beam-driven ion-cyclotron modes in the scrape-off layer of a field-reversed configuration",

\bibitem{Cottrell2000}
Cottrell G A 2000 \textit{Phys. Rev. Lett.} \textbf{84} 2397-400

\bibitem{Dendy2015}
Dendy R O and McClements K G 2015 \textit{Plasma Physics and Controlled Fusion} \textbf{57} 044002
%  "Ion cyclotron emission from fusion-born ions in large tokamak plasmas: a brief review from JET and TFTR to ITER",

\bibitem{McClements2015}
McClements K G, Dendy R O 2015 \textit{Nucl. Fusion} \textbf{55} 043013
% , "Fast particle-driven ion cyclotron emission (ICE) in tokamak plasmas and the case for an ICE diagnostic in ITER",

\bibitem{Belikov1976}
Belikov V S and Kolesnichenko Ya I 1976 \textit{Sov. Phys. Tech. Phys.} \textbf{20} 1146

\bibitem{Stix}
Stix T H 1992 \textit{Waves in Plasmas. 2nd Edition} (Springer, New York)

\bibitem{Dendy1994}
Dendy R O, Lashmore-Davies C N, McClements K G and Cottrell G A 1994 \textit{Physics of Plasmas} \textbf{1} 1918-28
%  "The excitation of obliquely propagating
% fast Alfvén waves at fusion ion cyclotron
% harmonics"

\bibitem{McClements1996}
McClements K G, Dendy R O, Lashmore-Davies C N, Cottrell G A, Cauffman S and Majeski R (1996) \textit{Phys. Plasmas} \textbf{3} 543

\bibitem{Cook2013}
Cook J W S, Dendy R O, Chapman S C 2013 \textit{Plasma Physics and Controlled Fusion} \textbf{55} 065003
% , "Particle-in-cell simulations of the magnetoacoustic cyclotron instability of fusion-born alpha-particles in tokamak plasmas",

\bibitem{Carbajal2014}
Carbajal L, Dendy R O, Chapman S C, and Cook J W S 2014 \textit{Phys. Plasmas} \textbf{21} 012106
% , "Linear and non-linear physics of the magnetoacoustic cyclotron instability of fusion-born ions in relation to ion cyclotron emission,"

\bibitem{Reman2019}
Reman B, Dendy R O, Akiyama T, Chapman S C, Cook J W S, Igami  H, Inagaki S, Saito K, and Yun G, 2019 \textit{Nucl. Fusion} \textbf{59} 096013
% “Interpreting observations of ion cyclotron emission from large helical device plasmas with beam-injected ion populations,”

\bibitem{Carbajal2021}
Carbajal L, and Calderón F A 2021 \textit{Phys. Plasmas} \textbf{28} 014505
% , "On the 2D dynamics of the magnetoacoustic cyclotron instability driven by fusion-born ions",

\bibitem{Sharapov2014}
Sharapov S E , Lilley M K, Akers R, Ayed N B, Cecconello M, Cook J W S, Cunningham G, and Verwichte E 2014 \textit{Phys. Plasmas} \textbf{21} 082501
% , "Bi-directional Alfvén cyclotron instabilities in the mega-amp spherical tokamak"

\bibitem{Gorelenkov2016}
Gorelenkov N N 2016 \textit{New J. Phys.} \textbf{18} 105010

\bibitem{Kolesnichenko2000}
Kolesnichenkoo Ya I, F{\"u}l{\"o}p T, Lisak M and Anderson D 1998 \textit{Nucl. Fusion} {\textbf 38} 1871
% , "Localized fast magnetoacoustic eigenmodes in tokamak plasmas"

\bibitem{Fulop2000}
F{\"u}l{\"o}p T, Lisak M, Kolesnichenko Ya I, and Anderson D, 2000 \textit{Phys. Plasmas} \textbf{7} 1479–1486
% “The radial and poloidal localization of fast magnetoacoustic eigenmodes in tokamaks,”

\bibitem{Heidbrink2006}
Heidbrink W W, Fredrickson E, Gorelenkov N N, Rhodes T, Zeeland  M A 2006 \textit{Nucl. Fusion} \textbf{46} 324
% Observation of compressional Alfvén eigenmodes (CAE) in a conventional tokamak.

\bibitem{Bezanson2017}
Bezanson J, Edelman A, Karpinski S, and Shah V B 2017 \textit{SIAM review} \textbf{59} 1
% Julia: A fresh approach to numerical computing,

\bibitem{Fredrickson2019}
Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
% "Emission in the ion cyclotron range of frequencies (ICE) on NSTX and NSTX-U",

\bibitem{Kronrod1965}
Kronrod A S 1965 \textit{Nodes and weights of quadrature formulas. Sixteen-place tables} (New York: Consultants Bureau (Authorized translation from the Russian))

\bibitem{QuadGK}
Johnson S G \textit{QuadGK.jl: Gauss-Kronrod integration in Julia} https://github.com/JuliaMath/QuadGK.jl

\bibitem{Nelder1965}
Nelder J A and Mead R 1965 \textit{Computer Journal} \textbf{7} 4
% , A simplex method for function minimization,

% \bibitem{Brambilla}
% M. Brambilla, Kinetic Theory of Plasma Waves: Homogeneous Plasmas (International Series of Monographs on Physics (96))

\bibitem{Fried1961}
Fried B D and Conte S C 1961 \textit{The Plasma Dispersion Function} (Academic Press, New York)

\bibitem{Sampoorna2007}
Sampoorna M, Nagendraa K N, Frisch H 2007 \textit{Journal of Quantitative Spectroscopy \& Radiative Transfer} \textbf{104} 71-85

\end{thebibliography}

\end{document}
% Splitting of spectral peaks
% ‘G. A. Cottrell, V. P. Bhatnagar, 0. da Costa, R. 0. Dendy, A. Edwards, J. Jacquinot, F. Nave, M. Schmid, A. Sibley, P. Smeulders, and D. F. H. Start, in Proceedings of the 19th European Conference on Controlled Fusion and Plasma Heating, Innsbruck, Austria, 1992, edited by W. Freysinger, K. Lackner, R. Schrittwieser, and W. Lindinger (European Physical Soci- ety, Petit-Lancy, Switzerland, 1990). Part I, Vol. 16C. p. 327.
% ‘The JET Team, Nucl. Fusion 32, 187 (1992).
% 7C. A. Cottrell, V. P. Bhatnagar, 0. da Costa, R. 0. Dendy, J. Jacquinot, K.
% G. McClements, D. C. McCune, M. E F. Nave, P. Smeulders, and D. F. H.
% Start, Nucl. Fusion 33, 1365 (1993).

### 2022-01-25 manuscript.tex
reproducible paper build

@@
 \documentclass[12pt]{iopart}
+\expandafter\let\csname equation*\endcsname\relax
+\expandafter\let\csname endequation*\endcsname\relax
 \usepackage{amsmath}
@@
 Solutions to the linearised Maxwell-Vlasov system of equations are found in $(k_\parallel, k_\bot)$ space for a ring-beam distribution of energetic minority alpha-particles and Maxwellian electrons and background thermal majority deuteron ions.
-Calculations are made for alpha-particles with ring-beams of two pitch values: $u_\parallel=-0.64 v_0 \simeq -1.085 V_A$, which is associated with the ion-cyclotron emission observed on JET Pulse No.
+Calculations are made for alpha-particles with ring-beams of two pitch-angle cosine values: $u_\parallel=-0.64 u_0 \simeq -1.085 V_A$, which is associated with the ion-cyclotron emission observed on JET Pulse No.
 26148, see Figs.
 3 and 15 and associated text in Ref.
-\cite{Cottrell1993}; and $u_\parallel=0 v_0$ representing a ring distribution with no parallel drift, see Fig.
+\cite{Cottrell1993}; and $u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see Fig.
 1(a) of Ref \cite{Cook2013}.
-Here $v_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is the speed of an alpha-particle with an energy of 3.6 MeV.
+Here $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is the speed of an alpha-particle with an energy of 3.6 MeV.
 The complex frequency solutions, $\omega$, are found corresponding to the fast Alfv{\'e}n branch on a grid of $(k_\parallel, k_\bot)$.
@@
 The component parallel (perpendicular) to the magnetic field is a drifting Maxwellian with thermal speed $v_{\parallel, th}$ ($v_{\perp, th}$) and drift velocity $u_\parallel$ ($u_\perp$).
-The thermal spread of the alpha particles' distribution function $v_{th}=v_{\parallel,th}=v_{\perp,th}=v_0/100$.
+The thermal spread of the alpha particles' distribution function $v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$.
 The concentration ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.
@@
 
-Solutions to the linearised Maxwell-Vlasov system of equations are given by the dispersion relation, Eq A.1.
-The hot dielectric tensor\cite{Stix} given by Eq. A.2 is calculated for each species.
+Solutions to the linearised Maxwell-Vlasov system of equations are given by the dispersion relation, Eq.
+\ref{eq:dettensor}, which is satisfied for sets of $(\omega,\vec{k})$.
+The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$,  given by Eq.
+\ref{eq:dielectrictensor} is calculated for each species.
+
+\begin{align}
+    0 &= ||\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}||,\label{eq:dettensor}\\
+    \vec{\vec{\epsilon}} &= \vec{\vec{1}} + \sum_s \frac{\Pi_s^2}{\omega}\sum_{n=-\infty}^{\infty}
+    \int_0^{\infty}
+    \int_{-\infty}^{\infty}\frac{\vec{\vec{S}}_{s,n} 2\pi v_\perp dv_\perp dv_\parallel}{\omega - k_\parallel v_\parallel - n \Omega_s}
+    \label{eq:dielectrictensor}
+\end{align}
+
+\noindent where
+
+\begin{align}
+\vec{\vec{S}}_{s,n} &=
+\begin{bmatrix}
+\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n J_n^2}{z_s} v_\perp W_s \\
+\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp W_s \\
+\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2 v_\parallel W_s \\
+\end{bmatrix}\label{eq:stenor},\\
+U_s &= \frac{\partial f_s}{\partial v_\perp} + \frac{k_\parallel}{\omega}  \left(v_\perp \frac{\partial f_s}{\partial v_\parallel} - v_\parallel \frac{\partial f_s}{\partial v_\perp}\right)\label{eq:uterm},\\
+W_s &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp}\label{eq:wterm},
+\end{align}
+
+\noindent $J_n$ is the $n$th Bessel function of the first kind and takes the argument $z_s=\frac{k_\perp v_\perp}{\Omega_s}$ where $J_n'=\frac{\partial J_n}{\partial z}$.
 
@@
 
-Since all the parallel distribution functions for all species are Maxwellians, the parallel integrals can be calculated from the well known plasma dispersion function\cite{Fried1961}, Eqs. B.1-B.3.
+Since all the parallel distribution functions for all species are Maxwellians, the parallel integrals can be calculated from the well known plasma dispersion function\cite{Fried1961}, which is given in the appendix.
 
@@
 Root finding of Eq.
-A.1 is performed by an amalgamation of the Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method, whereby if the Nelder-Mead simplex does not contain a root, as calculated via the winding method, then the optimisation procedure follows the Nelder-Mead method, otherwise if a root is present with the simplex then the simplex bifurcates and the child simplex that contains the root is used for the next iteration.
+\ref{eq:dettensor} is performed by an amalgamation of the Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method, whereby if the Nelder-Mead simplex does not contain a root, as calculated via the winding method, then the optimisation procedure follows the Nelder-Mead method, otherwise if a root is present with the simplex then the simplex bifurcates and the child simplex that contains the root is used for the next iteration.
 The optimisation iteration procedure finishes when the real extent of the simplex is less than $10^{-4}$ of the alpha-particle cyclotron frequency $\Omega_i$ (the only ion species considered are deuterons and alpha-particles, which have identical cyclotron frequencies) and the imaginary extent is less than $10^{-5}\,\Omega_i$.
@@
 
-The growth rates for the zero-pitch case ($u_\parallel = 0 v_0$ in Eq.
+The growth rates of unstable modes for the case where the pitch-angle cosine value is zero ($u_\parallel = 0$ in Eq.
 1) are plotted in Fig.
-\ref{fig:2D_F12_zeropitch} against the associated real frequency (damped modes are not shown).
-In this figure shading indicates the propagation angle, where $90^\circ$ ($0^\circ$) is exactly perpendicular (parallel) to the magnetic field.
+\ref{fig:2D_F12_zeropitch}(a) against the associated real frequency (only modes with growth rates in excess of the solver tolerance are shown i.e.
+$\Im(\omega) > 10^{-5} \Omega_i$).
+In this panel shading indicates the propagation angle, where $90^\circ$ ($0^\circ$) is exactly perpendicular (parallel) to the magnetic field.
+Fig.
+\ref{fig:2D_F12_zeropitch}(b) plots the same data in the $(k_\parallel,\omega)$ plane where shading instead indicates growth rates.
+The simplicity of the consecutive unstable cyclotron harmonics visible in panel (a) betrays the complexity of the structure in the $(k_\parallel, \omega)$ plane.
+Integration over $k_\parallel$ projects the solutions onto the frequency axis, which loses the structure and produces narrow harmonics.
 Fig.
 \ref{fig:2D_imag_zeropitch} plots shows the same data as Fig.
-\ref{fig:2D_F12_zeropitch} except shading indicates the growth rates in $(k_\parallel, k_\perp)$ space.
+\ref{fig:2D_F12_zeropitch} with damped and normal modes in addition, except shading indicates the growth rates in $(k_\parallel, k_\perp)$ space.
 Overlaid contours indicate the real frequency in units of ion cyclotron harmonics.
@@
 Together these figures serves to introduce a 2D view (growth rates in $(k_\parallel, k_\perp)$ space) of traditionally 1D data (growth rates against frequency).
-This will become important when a non-zero pitch is taken into account next.
+This will become important when a non-zero $u_\parallel$ is taken into account next.
 
 \begin{figure}[ht!]
-    \centering
-\includegraphics[scale=0.8]{ICE2D_F12_zeropitch.pdf}
-    \caption{(colour online) The imaginary component of the frequencies of the unstable fast Alfv{\'e}n wave solutions to the linear Maxwell-Vlasov system of equations, Eq.
-A.1, for thermal electrons, deuterons and a ring distribution of energetic alpha-particles i.e.
+    % \centering
+    \raggedleft
+\includegraphics[scale=0.9]{fig1.pdf}
+    \caption{(colour online) Unstable fast Alfv{\'e}n wave solutions to the linear Maxwell-Vlasov system of equations, Eq.
+\ref{eq:dettensor}, for thermal electrons, deuterons and a ring distribution of energetic alpha-particles i.e.
 when there is zero parallel drift, $u_\parallel=0$ in Eq.
 1.
+(a) Upper panel: growth rate as a function of frequency where shading indicates the propagation angle of the complex frequency solution.
 Consecutive cyclotron harmonics are increasingly unstable and are narrow due to the lack of Doppler shift associated with parallel drifts.
-Shading indicates the propagation angle of the complex frequency solution.
-These data are also shown in Fig.
-\ref{fig:2D_imag_zeropitch} in $(k_\parallel,k_\perp)$ space.
-\ref{fig:2D_real_zeropitch}.}
+(b) Lower panel: the same solutions as in panel (a) in the $(k_\parallel,\omega)$ plane where shading indicates growth rate.
+The complex structure of the growth rates that exists in wavenumber space is projected into sharp harmonics on the frequency axis.}
     \label{fig:2D_F12_zeropitch}
@@
 \begin{figure}[ht!]
-    \centering
-\includegraphics[scale=0.8]{ICE2D_imag_zeropitch.pdf}
+    % \centering
+    \raggedleft
+\includegraphics[scale=0.9]{fig2.pdf}
     \caption{(colour online) Shading indicates the growth rate of fast Alfv{\'e}n waves excited by a ring of fusion born alpha-particles, as a function of parallel wavenumber on the ordinate and perpendicular wavenumber on the abscissa.
@@
 Curved grey traces indicate the location in $(k_\parallel, k_\perp)$ of the cyclotron harmonics (i.e.
-the real counterpart to the imaginary frequency), which are annotated along the left hand side and upper edge of the figure.
+the real component of the complex $\omega$ solutions), which are annotated along the left hand side and upper edge of the figure.
 Frequencies are in units of the alpha-particle cyclotron frequency, and wavenumbers are in alpha-particle cyclotron frequencies per Alfv{\'e}n speed.
@@
 Fig.
-\ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch case ($u_\parallel = -0.64 v_0$) in the same manner as Fig.
+\ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch-angle case ($u_\parallel = -0.64 u_0$) in the same manner as Fig.
 \ref{fig:2D_F12_zeropitch}.
@@
 Second, harmonics are broader due to the non-zero parallel drift that gives rise to strong Doppler shift with increasing $|k_\parallel|$.
-Third, doublet splitting of the $8^{th}$ harmonic, for example, is observed as it is in Fig.
-2 of Cottrell 1993 and as is indicated by Fig.
+Third, doublet splitting of the $10^{th}$ harmonic is observed: see Fig.
+2 of Cottrell 1993 for a single signal showing multiple doublet splittings; Fig.
 3 of Dendy et al 1994 (also Fig.
-9 Dendy et al 1995), which shows slices at constant propagation angle that when all superimposed point towards doublet splitting, albeit under-resolved in wavenumber space.
+9 Dendy et al 1995) shows slices at constant propagation angle that when all superimposed point towards doublet splitting, albeit under-resolved in wavenumber space.
 The following describes how these three features are connected.
@@
 The effect is exaggerated for higher harmonics where the growth rate diminishes to zero for perpendicular propagation; see e.g.
-the $10^{th}$ harmonic in Fig.
-\ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A, k_\bot \simeq 11\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron frequency and $V_A$ is the Alfv{\'e}n speed.
+the $12^{th}$ harmonic in Fig.
+\ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A, k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron frequency and $V_A$ is the Alfv{\'e}n speed.
 Other possible mechanisms for doublet splitting could be grad-B or curvature drifts\cite{Cottrell1993}.
 Qualitatively, Fig.
-\ref{fig:2D_F12_physicalpitch} resembles Fig.
+\ref{fig:2D_F12_physicalpitch}(a) resembles Fig.
 2 of Cottrell 1993 in that the signal below $\sim 7$ cyclotron harmonics is lower and flatter with well defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change larger and the harmonics are less easily identifiable.
@@
 Modes propagating exactly perpendicular become stable in the region $k_\perp > 13 \Omega_i/V_A$ whilst obliquely propagating modes remain unstable; see Fig.
-9 of Dendy et al 1995.
+9 of Dendy et al 1995, which plots the analytic expression for the growth rate of the magnetoacoustic cylcotron instability, i.e.
+the instability thought responsible for ICE, at 5 slices of propagation angle.
 The Doppler shift of the growing modes can reach $\approx \Omega_i$ for large propagation angles at lower frequencies and also for moderately oblique angles at higher frequencies; see, for example, the mode associated with the $15^{th}$ cyclotron harmonic at, $(k_\parallel \simeq -\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e.
@@
 \begin{figure}[ht!]
-    \centering
-\includegraphics[scale=0.8]{ICE2D_F12_physicalpitch.pdf}
-    \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 v_0$.
-See also Fig.
-\ref{fig:2D_imag_physicalpitch} plotted against the real part, shown in \ref{fig:2D_real_zeropitch}.
-Shading indicates the wave's propagation angle.
-Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks.
-The 8th cyclotron harmonic is split into a doublet, or has two peaks because both the growth rate and real frequency vary with propagation angle.
+    % \centering
+    \raggedleft
+\includegraphics[scale=0.9]{fig3.pdf}
+    \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
+Panel (a) plots the growth rate against frequency where shading indicates the wave's propagation angle, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
+Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks in frequency.
+The 10th cyclotron harmonic is split into a doublet, or has two peaks because both the growth rate and real frequency vary with wavenumber.
 Compare with Fig.
@@
 \begin{figure}[ht!]
-    \centering
-\includegraphics[scale=0.8]{ICE2D_imag_physicalpitch.pdf}
+    % \centering
+    \raggedleft
+\includegraphics[scale=0.9]{fig4.pdf}
     \caption{(colour online) As in Fig.
-\ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch value resulting in $u_{\parallel} = -0.64 v_0$.
-The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch of the opposite sign, $u_{\parallel} = 0.64 v_0$, is identical expect flipped about $k_\parallel=0$.}
+\ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch-angle value resulting in $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
+The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch-angle cosine of the opposite sign, $u_{\parallel} = 0.64 u_0$, is identical expect flipped about $k_\parallel=0$.}
     \label{fig:2D_imag_physicalpitch}
@@
 \begin{figure}[ht!]
-    \centering
-\includegraphics[scale=0.8]{ICE2D_real_div_guess_zeropitch.pdf}
+    % \centering
+    \raggedleft
+\includegraphics[scale=0.9]{fig5.pdf}
     \caption{(colour online) As in Fig.
@@
 \cite{McClements1996}.
-The differences in real frequency for the two sets of calculations, physical pitch of $-0.64$ and zero pitch, are negligible.}
+The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.64$ and zero pitch, are negligible.}
     \label{fig:2D_real_zeropitch}
@@
 \appendix
-\section{}
-
-The plasma dispersion relation\cite{Stix} is given by the $(\omega,\vec{k})$ solutions to A.1, given the contributions of the dielectric tensor $\vec{\vec{\epsilon}}$ from all species.
-
-\begin{align}
-    0 &= ||\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}||,\\
-    \vec{\vec{\epsilon}} &= \vec{\vec{1}} + \sum_s \frac{\Pi_s^2}{\omega}\sum_{n=-\infty}^{\infty}
-    \int_0^{\infty}
-    \int_{-\infty}^{\infty}\frac{\vec{\vec{S}}_{s,n} 2\pi v_\perp dv_\perp dv_\parallel}{\omega - k_\parallel v_\parallel - n \Omega_s}
-\end{align}
-
-where
-
-\begin{align}
-\vec{\vec{S}}_{s,n} &=
-\begin{bmatrix}
-\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n J_n^2}{z_s} v_\perp W_s\\
-\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp W_s \\
-\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2 v_\parallel W_s \\
-\end{bmatrix},\\
-U_s &= \frac{\partial f_s}{\partial v_\perp} + \frac{k_\parallel}{\omega}  \left(v_\perp \frac{\partial f_s}{\partial v_\parallel} - v_\parallel \frac{\partial f_s}{\partial v_\perp}\right), \\
-W_s &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp},\\
-\end{align}
-
-$J_n$ is the $n$th Bessel function of the first kind and takes the argument $z_s=\frac{k_\perp v_\perp}{\Omega_s}$ where $J_n'=\frac{\partial J_n}{\partial z}$.
 
@@
 
-can be calculated via the recurrence relation\cite{Sampoorna2007}
+\noindent can be calculated via the recurrence relation\cite{Sampoorna2007}
 
@@
 
-where $n \geq 0$ and the cases $n=1$ and $n=3$ are stated explicitly to avoid confusion. The zeroth term is defined as
+\noindent where $n \geq 0$ and the cases $n=1$ and $n=3$ are stated explicitly to avoid confusion. The zeroth term is defined as
 
@@
 
-where $\mathrm{erfcx}$ is the scaled complementary complex error function.
+\noindent where $\mathrm{erfcx}$ is the scaled complementary complex error function.
 

### 2022-01-25 ICE2D.jl

@@ -1,0 +1,480 @@
+using Distributed, Dates
+using Plots, Random, ImageFiltering, Statistics
+using Dierckx, Contour, JLD2
+
+println("Starting at ", now())
+# pitchanglecosine of 0 is all vperp, and (-)1 is (anti-)parallel to B field
+const pitchanglecosine = try; parse(Float64, ARGS[1]); catch; -0.64; end
+@assert -1 <= pitchanglecosine <= 1
+# thermal width of ring as a fraction of its speed # Dendy PRL 1993
+const vthermalfractionz = try; parse(Float64, ARGS[2]); catch; 0.01; end
+const vthermalfraction⊥ = try; parse(Float64, ARGS[3]); catch; 0.01; end
+const name_extension = if length(ARGS) >= 4
+  ARGS[4]
+else
+  "$(pitchanglecosine)_$(vthermalfractionz)_$(vthermalfraction⊥)"
+end
+const filecontents = [i for i in readlines(open(@__FILE__))]
+const nprocsadded = div(Sys.CPU_THREADS, 2)
+
+addprocs(nprocsadded)
+
+@everywhere using ProgressMeter # for some reason must be up here on its own
+@everywhere using StaticArrays
+@everywhere using FastClosures
+@everywhere begin
+  using PlasmaDispersionRelations, LinearAlgebra, WindingNelderMead
+
+  mₑ = PlasmaDispersionRelations.mₑ
+  md = 2*1836*mₑ
+  mα = 2*md
+
+  # Fig 18 Cottrell 1993
+  n0 = 1.7e19 # central electron density 3.6e19
+  B0 = 2.07 # OR 2.191 for B field at 4m when 2.8 T on axis R0 3.13m
+  ξ = 1.5e-4 # nα / ni = 1.5 x 10^-4
+
+  nd = n0 / (1.0 + 2*ξ)
+  nα = ξ*nd
+  @assert n0 ≈ 2*nα + nd
+  Va = sqrt(B0^2/PlasmaDispersionRelations.μ₀/nd/md)
+
+  Ωe = cyclotronfrequency(B0, mₑ, -1)
+  Ωd = cyclotronfrequency(B0, md, 1)
+  Ωα = cyclotronfrequency(B0, mα, 2)
+  Πe = plasmafrequency(n0, mₑ, -1)
+  Πd = plasmafrequency(nd, md, 1)
+  Πα = plasmafrequency(nα, mα, 2)
[…]

### 2022-01-25 run.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+source ~/.bashrc
+shopt -s expand_aliases
+
+# first arg is pitch, second arg is ring thermal spread as fraction of its speed
+julia1.7 -e 'using Pkg; Pkg.instantiate()'
+julia1.7 --proj ICE2D.jl -0.64 0.01 0.01 physical
+julia1.7 --proj ICE2D.jl  0.0  0.01 0.01 ring
+
+for j in physical ring
+do
+  for i in `ls *$j.pdf`
+  do
+    pdf2ps $i $i.ps
+    ps2pdf $i.ps $i
+    rm $i.ps
+  done
+done
+
+cp ICE2D_Combo_ring.pdf fig1.pdf
+cp ICE2D_imag_ring.pdf fig2.pdf
+cp ICE2D_Combo_physical.pdf fig3.pdf
+cp ICE2D_imag_physical.pdf fig4.pdf
+cp ICE2D_real_div_guess_ring.pdf fig5.pdf
+
+pdflatex manuscript.tex
+pdflatex manuscript.tex

### 2022-01-31 manuscript.tex
making improvements

@@
 
-Fig.
-\ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch-angle case ($u_\parallel = -0.64 u_0$) in the same manner as Fig.
-\ref{fig:2D_F12_zeropitch}.
-There are several points worth noting.
-First, the growth rate is no longer linear with cyclotron harmonic number.
-Second, harmonics are broader due to the non-zero parallel drift that gives rise to strong Doppler shift with increasing $|k_\parallel|$.
-Third, doublet splitting of the $10^{th}$ harmonic is observed: see Fig.
-2 of Cottrell 1993 for a single signal showing multiple doublet splittings; Fig.
-3 of Dendy et al 1994 (also Fig.
-9 Dendy et al 1995) shows slices at constant propagation angle that when all superimposed point towards doublet splitting, albeit under-resolved in wavenumber space.
-The following describes how these three features are connected.
-The doublet splitting arises due to the dip in growth rate where $k_\parallel$ goes through zero combined with the change in sign of the Doppler shift of the real frequency as $k_\parallel$ changes sign.
-The effect is exaggerated for higher harmonics where the growth rate diminishes to zero for perpendicular propagation; see e.g.
-the $12^{th}$ harmonic in Fig.
-\ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A, k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron frequency and $V_A$ is the Alfv{\'e}n speed.
-Other possible mechanisms for doublet splitting could be grad-B or curvature drifts\cite{Cottrell1993}.
+Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch-angle
+case ($u_\parallel = -0.64 u_0$) in the same manner as Fig.
+\ref{fig:2D_F12_zeropitch}. There are several points worth noting. First, the
+growth rate is no longer linear with cyclotron harmonic number. Second,
+harmonics are broader due to the non-zero parallel drift that gives rise to
+strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
+of the $10^{th}$ harmonic is observed: see Fig. 2 of Cottrell 1993 for a single
+signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
+Fig. 9 Dendy et al 1995) shows slices at constant propagation angle that when
+all superimposed point towards doublet splitting, albeit under-resolved in
+wavenumber space. The following describes how these three features are
+connected. The doublet splitting arises due to the dip in growth rate where
+$k_\parallel$ goes through zero combined with the change in sign of the Doppler
+shift of the real frequency as $k_\parallel$ changes sign. The effect is
+exaggerated for higher harmonics where the growth rate diminishes to zero for
+perpendicular propagation; see e.g. the $12^{th}$ harmonic in Fig.
+\ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
+k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron
+frequency and $V_A$ is the Alfv{\'e}n speed. Other possible mechanisms for
+doublet splitting could be grad-B or curvature
+drifts\cite{Cottrell1993,Fulop1998}.
 Qualitatively, Fig.
@@
 \cite{McClements1996}.
-The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.64$ and zero pitch, are negligible.}
+The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.64$ and zero, are negligible.}
     \label{fig:2D_real_zeropitch}
@@
 
-% \bibitem{Brambilla}
-% M. Brambilla, Kinetic Theory of Plasma Waves: Homogeneous Plasmas (International Series of Monographs on Physics (96))
+\bibitem{Fulop1998}
+F{\"u}l{\"o}p T and Lisak M 1998 \textit{Nucl. Fusion} \textbf{38} 761
 

### 2022-01-31 ICE2D.jl

@@ -50,7 +50,7 @@ addprocs(nprocsadded)
   vα = thermalspeed(3.6e6, mα)
   # pitchanglecosine = cos(pitchangle)
   # acos(pitchanglecosine) = pitchangle
-  pitchanglecosine = Float64(@fetchfrom 1 pitchanglecosine) 
+  pitchanglecosine = Float64(@fetchfrom 1 pitchanglecosine)
   vα⊥ = vα * sqrt(1 - pitchanglecosine^2) # perp speed
   vαz = vα * pitchanglecosine # parallel speed
   vthermalfractionz = Float64(@fetchfrom 1 vthermalfractionz)
@@ -74,18 +74,9 @@ addprocs(nprocsadded)
   alpha_delta = SeparableVelocitySpecies(Πα, Ωα,
     FParallelDiracDelta(vαz),
     FPerpendicularDiracDelta(vα⊥))
-  alpha_beamdelta = SeparableVelocitySpecies(Πα, Ωα,
-    FBeam(vαthz, vαz),
-    FPerpendicularDiracDelta(vα⊥))
-  alpha_deltaring = SeparableVelocitySpecies(Πα, Ωα,
-    FParallelDiracDelta(vαz),
-    FRing(vαth⊥, vα⊥))
-  alpha_shell = CoupledVelocitySpecies(Πα, Ωα,
-    FShell(sqrt(vαthz^2 + vαth⊥^2), vα))
 
   Smmr = Plasma([electron_maxw, deuteron_maxw, alpha_ringbeam])
   Smmd = Plasma([electron_maxw, deuteron_maxw, alpha_delta])
-  #Smmb = Plasma([electron_maxw, deuteron_maxw, alpha_beamdelta])
 
   f0 = abs(Ωα)
   k0 = f0 / abs(Va)
@@ -313,8 +304,6 @@ function plotit(sols, file_extension=name_extension, fontsize=9)
   end
   xlabel = "\$\\mathrm{Perpendicular\\ Wavenumber} \\ [\\Omega_{i} / V_A]\$"
   ylabel = "\$\\mathrm{Parallel\\ Wavenumber} \\ [\\Omega_{i} / V_A]\$"
-  #colorgrad = Plots.cgrad([:cyan, :lightblue, :blue, :darkblue, :black,
-  #                        :darkred, :red, :orange, :yellow])
   zs = real.(ωs)
   climmax = maximum(zs)
   plotter2d(zs, xlabel, ylabel, Plots.cgrad(), 0.0, climmax)
[…]

### 2022-01-31 manuscript.tex
quantisation effect of toroidicity

@@
 
-The growth rates of unstable modes for the case where the pitch-angle cosine value is zero ($u_\parallel = 0$ in Eq.
-1) are plotted in Fig.
-\ref{fig:2D_F12_zeropitch}(a) against the associated real frequency (only modes with growth rates in excess of the solver tolerance are shown i.e.
-$\Im(\omega) > 10^{-5} \Omega_i$).
-In this panel shading indicates the propagation angle, where $90^\circ$ ($0^\circ$) is exactly perpendicular (parallel) to the magnetic field.
+The growth rates of unstable modes for the case where the pitch-angle cosine
+value is zero ($u_\parallel = 0$ in Eq. 1) are plotted in Fig.
+\ref{fig:2D_F12_zeropitch}(a) against the associated real frequency (only modes
+with growth rates in excess of the solver tolerance are shown i.e. $\Im(\omega)
+> 10^{-5} \Omega_i$). In this panel shading indicates the parallel wavenumber in
+units of $\Omega_i/V_A$.
 Fig.
@@
 \includegraphics[scale=0.9]{fig1.pdf}
-    \caption{(colour online) Unstable fast Alfv{\'e}n wave solutions to the linear Maxwell-Vlasov system of equations, Eq.
-\ref{eq:dettensor}, for thermal electrons, deuterons and a ring distribution of energetic alpha-particles i.e.
-when there is zero parallel drift, $u_\parallel=0$ in Eq.
-1.
-(a) Upper panel: growth rate as a function of frequency where shading indicates the propagation angle of the complex frequency solution.
+    \caption{(colour online) Unstable fast Alfv{\'e}n wave solutions to the
+    linear Maxwell-Vlasov system of equations, Eq. \ref{eq:dettensor}, for
+    thermal electrons, deuterons and a ring distribution of energetic
+    alpha-particles i.e. when there is zero parallel drift, $u_\parallel=0$ in
+    Eq. 1. (a) Upper panel: growth rate as a function of frequency where shading
+    indicates the parallel wavenumber in units of $\Omega_i/V_A$ of the complex frequency solution.
 Consecutive cyclotron harmonics are increasingly unstable and are narrow due to the lack of Doppler shift associated with parallel drifts.
@@
     \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
-Panel (a) plots the growth rate against frequency where shading indicates the wave's propagation angle, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
+Panel (a) plots the growth rate against frequency where shading indicates the wave's parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
 Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks in frequency.
@@
 
-These calculations capture the finite perpendicular thermal spread of the alpha particle distribution function, the integral of which has no known analytically tractable form.
-The computational capabilities of the software, PlasmaDispersionRelations.jl, used to generate the results further enable ICE as a diagnostic for fusion plasmas since the code is capable of: calculating the dispersion relation of plasmas consisting of an arbitrary number of species; and of arbitrary distribution functions, although those with parallel Maxwellians, like those presented here, are faster.
-It has been shown that it is possible to recreate signals containing lower harmonics of substantial amplitudes by invoking 2D linear physics.
-The presence of ICE signals at lower harmonics in PIC simulations has been attributed to nonlinear interactions between pairs of more powerful higher harmonics\cite{Carbajal2014}.
-This work suggests that observations of lower ICE harmonics may arise either due to the effects of 2D linear physics or to nonlinear 1D physics\cite{Chapman2018}, or a combination.
-An important feature missing from these calculations is the effect of plasma geometry, which impacts the ICE signals according to frequency and propagation angle.
+These calculations capture the finite perpendicular thermal spread of the alpha
+particle distribution function, the integral of which has no known analytically
+tractable form. The computational capabilities of the software,
+PlasmaDispersionRelations.jl, used to generate the results further enable ICE as
+a diagnostic for fusion plasmas since the code is capable of: calculating the
+dispersion relation of plasmas consisting of an arbitrary number of species; and
+of arbitrary distribution functions, although those with parallel Maxwellians,
+like those presented here, are faster. It has been shown that it is possible to
+recreate signals containing lower harmonics of substantial amplitudes by
+invoking 2D linear physics. The presence of ICE signals at lower harmonics in
+PIC simulations has been attributed to nonlinear interactions between pairs of
+more powerful higher harmonics\cite{Carbajal2014}. This work suggests that
+observations of lower ICE harmonics may arise either due to the effects of 2D
+linear physics or to nonlinear 1D physics\cite{Chapman2018}, or a combination.
+There are several important features missing from these calculations. Among them
+are the effect the spatial inhomogeniety of the magnetic field and background
+plasma, and the quantisation associatied with toroidal periodicity. These may
+create an eigenmode structure into which the MCI must fit.
 Cottrell et al.

### 2022-01-31 manuscript.tex
tiny word change in caption

@@
     \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
-Panel (a) plots the growth rate against frequency where shading indicates the wave's parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
+Panel (a) plots growth rate against frequency where shading indicates parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
 Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks in frequency.

### 2022-02-02 manuscript.tex
tiny word change in caption
updates to calc of pitch -0.646

@@
 
-Solutions to the linearised Maxwell-Vlasov system of equations are found in $(k_\parallel, k_\bot)$ space for a ring-beam distribution of energetic minority alpha-particles and Maxwellian electrons and background thermal majority deuteron ions.
-Calculations are made for alpha-particles with ring-beams of two pitch-angle cosine values: $u_\parallel=-0.64 u_0 \simeq -1.085 V_A$, which is associated with the ion-cyclotron emission observed on JET Pulse No.
-26148, see Figs.
-3 and 15 and associated text in Ref.
-\cite{Cottrell1993}; and $u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see Fig.
+Solutions to the linearised Maxwell-Vlasov system of equations are found in
+$(k_\parallel, k_\bot)$ space for a ring-beam distribution of energetic minority
+alpha-particles and Maxwellian electrons and background thermal majority
+deuteron ions. Calculations are made for alpha-particles with ring-beams of two
+pitch-angle cosine values: $u_\parallel=-0.646 u_0 \approx
+-\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_1})) \simeq -1.1 V_A$, which is
+associated with the ion-cyclotron emission observed on JET Pulse No. 26148 for
+which $R_0=3.46m$ and
+$\psi_0=55.2^\circ$ are the major radius and angle pitch angle at birth, and
+$R_1=4m$ is the major radius of the emission region, see
+Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
+$u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
+Fig.
 1(a) of Ref \cite{Cook2013}.
@@
 
-The plasma parameters approximate those found in the region indicated in Fig.
-3 of Ref.
-\cite{Cottrell1993}, which are associated with the continuous trace signal of Fig.
-2 (ibid).
-The electrons and deuterons, both modelled as Maxwellians, have temperatures of 1 keV and the electrons have a number density of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$.
-The magnetic field is $2.07$ $\mathrm{T}$.
+The plasma parameters approximate those found in the region indicated in Fig. 3
+of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
+signal of Fig. 2 (ibid). The electrons and deuterons, both modelled as
+Maxwellians, have temperatures of 1 keV and the electrons have a number density
+of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
+$\mathrm{T}$, which corresponds to a deuterium cyclotron frequency of $17MHz$.
 The ring-beam distribution function, Eq.
@@
 \begin{bmatrix}
-\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n J_n^2}{z_s} v_\perp W_s \\
-\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp W_s \\
-\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2 v_\parallel W_s \\
+\frac{n^2J_n^2}{z_s^2}v_\perp U_s & \frac{inJ_nJ_n'}{z_s}v_\perp U_s & \frac{n
+  J_n^2}{z_s} v_\perp W_{s,n} \\
+\frac{-inJ_nJ_n'}{z_s}v_\perp U_s & (J_n')^2 v_\perp U_s & -iJ_n J_n' v_\perp
+  W_{s,n} \\
+\frac{nJ_n^2}{z_s}v_\parallel U_s & iJ_nJ_n'v_\parallel U_s & n J_n^2
+  v_\parallel W_{s,n} \\
 \end{bmatrix}\label{eq:stenor},\\
 U_s &= \frac{\partial f_s}{\partial v_\perp} + \frac{k_\parallel}{\omega}  \left(v_\perp \frac{\partial f_s}{\partial v_\parallel} - v_\parallel \frac{\partial f_s}{\partial v_\perp}\right)\label{eq:uterm},\\
-W_s &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp}\label{eq:wterm},
+  W_{s,n} &=\left(1- \frac{n\Omega_s}{\omega}\right)\frac{\partial f_s}{\partial v_\parallel} + \frac{n\Omega v_\parallel}{\omega v_\perp}\frac{\partial f_s}{\partial v_\perp}\label{eq:wterm},
 \end{align}
@@
 with growth rates in excess of the solver tolerance are shown i.e. $\Im(\omega)
-> 10^{-5} \Omega_i$). In this panel shading indicates the parallel wavenumber in
+> 10^{-5} \Omega_i$). In this panel shading indicates parallel wavenumber in
 units of $\Omega_i/V_A$.
@@
 Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch-angle
-case ($u_\parallel = -0.64 u_0$) in the same manner as Fig.
+case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
 \ref{fig:2D_F12_zeropitch}. There are several points worth noting. First, the
@@
 strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
-of the $10^{th}$ harmonic is observed: see Fig. 2 of Cottrell 1993 for a single
+of the $11^{th}$ harmonic is observed: see Fig. 2 of Cottrell 1993 for a single
 signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
@@
 \includegraphics[scale=0.9]{fig3.pdf}
-    \caption{(colour online) The imaginary component of the frequencies for the case with parallel drift, $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
-Panel (a) plots growth rate against frequency where shading indicates parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions in the $k_\parallel, \omega$ plane where shading indicates growth rate.
-Note that the Doppler shift associated with the non-zero parallel drift speed broadens the unstable peaks in frequency.
-The 10th cyclotron harmonic is split into a doublet, or has two peaks because both the growth rate and real frequency vary with wavenumber.
-Compare with Fig.
-2 of Cottrel et al 1993.}
+    \caption{(colour online) The imaginary component of the frequencies for the
+    case with parallel drift, $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
+    Panel (a) plots growth rate against frequency where shading indicates
+    parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions
+    in the $k_\parallel, \omega$ plane where shading indicates growth rate. Note
+    that the Doppler shift associated with the non-zero parallel drift speed
+    broadens the unstable peaks in frequency. The $10^{th}$ and $11^{th}$ cyclotron
+    harmonics are split into doublets, i.e. they have two peaks because both
+    the growth rate and real frequency vary with wavenumber. Compare with
+    Fig. 2 of Cottrel et al 1993.}
     \label{fig:2D_F12_physicalpitch}
@@
     \caption{(colour online) As in Fig.
-\ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch-angle value resulting in $u_{\parallel} = -0.64 u_0 \simeq -1.085 V_A$.
-The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch-angle cosine of the opposite sign, $u_{\parallel} = 0.64 u_0$, is identical expect flipped about $k_\parallel=0$.}
+\ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch-angle value resulting in $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
+The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch-angle cosine of the opposite sign, $u_{\parallel} = 0.646 u_0$, is identical expect flipped about $k_\parallel=0$.}
     \label{fig:2D_imag_physicalpitch}
@@
 \cite{McClements1996}.
-The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.64$ and zero, are negligible.}
+The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.646$ and zero, are negligible.}
     \label{fig:2D_real_zeropitch}

### 2022-02-02 ICE2D.jl

@@ -1,10 +1,10 @@
 using Distributed, Dates
 using Plots, Random, ImageFiltering, Statistics
-using Dierckx, Contour, JLD2
+using Dierckx, Contour, JLD2, DelimitedFiles
 
 println("Starting at ", now())
 # pitchanglecosine of 0 is all vperp, and (-)1 is (anti-)parallel to B field
-const pitchanglecosine = try; parse(Float64, ARGS[1]); catch; -0.64; end
+const pitchanglecosine = try; parse(Float64, ARGS[1]); catch; -0.646; end
 @assert -1 <= pitchanglecosine <= 1
 # thermal width of ring as a fraction of its speed # Dendy PRL 1993
 const vthermalfractionz = try; parse(Float64, ARGS[2]); catch; 0.01; end
@@ -31,7 +31,8 @@ addprocs(nprocsadded)
 
   # Fig 18 Cottrell 1993
   n0 = 1.7e19 # central electron density 3.6e19
-  B0 = 2.07 # OR 2.191 for B field at 4m when 2.8 T on axis R0 3.13m
+  B0 = 2.07 # 2.07T = 2.8T * 2.96 m / 4m
+  # 2.23 T is 17MHz for deuterium cyclotron frequency
   ξ = 1.5e-4 # nα / ni = 1.5 x 10^-4
 
   nd = n0 / (1.0 + 2*ξ)
@@ -78,8 +79,8 @@ addprocs(nprocsadded)
   Smmr = Plasma([electron_maxw, deuteron_maxw, alpha_ringbeam])
   Smmd = Plasma([electron_maxw, deuteron_maxw, alpha_delta])
 
-  f0 = abs(Ωα)
-  k0 = f0 / abs(Va)
+  w0 = abs(Ωα)
+  k0 = w0 / abs(Va)
 
   γmax = abs(Ωα) * 0.15
   γmin = -abs(Ωα) * 0.075
@@ -105,9 +106,9 @@ addprocs(nprocsadded)
 
     config = Configuration(K, options)
 
-    ics = ((@SArray [ω0*0.8, f0*0.08]),
-           (@SArray [ω0*0.9, f0*0.04]),
-           (@SArray [ω0*1.0, f0*0.01]))
+    ics = ((@SArray [ω0*0.8, w0*0.08]),
+           (@SArray [ω0*0.9, w0*0.04]),
+           (@SArray [ω0*1.0, w0*0.01]))
 
 
[…]

### 2022-02-02 run.sh

@@ -4,9 +4,10 @@ shopt -s expand_aliases
 
 # first arg is pitch, second arg is ring thermal spread as fraction of its speed
 julia1.7 -e 'using Pkg; Pkg.instantiate()'
-julia1.7 --proj ICE2D.jl -0.64 0.01 0.01 physical
-julia1.7 --proj ICE2D.jl  0.0  0.01 0.01 ring
+julia1.7 --proj ICE2D.jl -0.646 0.01 0.01 physical
+julia1.7 --proj ICE2D.jl  0.0   0.01 0.01 ring
 
+# now reduce size and improve images
 for j in physical ring
 do
   for i in `ls *$j.pdf`

### 2022-02-05 manuscript.tex
minor updates to manuscript

@@
 of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
-$\mathrm{T}$, which corresponds to a deuterium cyclotron frequency of $17MHz$.
+$\mathrm{T}$, which corresponds a major radius of $4m$.
 The ring-beam distribution function, Eq.
@@
 > 10^{-5} \Omega_i$). In this panel shading indicates parallel wavenumber in
-units of $\Omega_i/V_A$.
-Fig.
-\ref{fig:2D_F12_zeropitch}(b) plots the same data in the $(k_\parallel,\omega)$ plane where shading instead indicates growth rates.
-The simplicity of the consecutive unstable cyclotron harmonics visible in panel (a) betrays the complexity of the structure in the $(k_\parallel, \omega)$ plane.
+units of $\Omega_i/V_A$. Fig. \ref{fig:2D_F12_zeropitch}(b) plots the same data
+in the $(\omega,k_\parallel)$ plane where shading instead indicates growth
+rates. The simplicity of the consecutive unstable cyclotron harmonics visible in
+panel (a) betrays the complexity of the structure in the $(\omega,k_\parallel)$ plane.
 Integration over $k_\parallel$ projects the solutions onto the frequency axis, which loses the structure and produces narrow harmonics.
@@
     Eq. 1. (a) Upper panel: growth rate as a function of frequency where shading
-    indicates the parallel wavenumber in units of $\Omega_i/V_A$ of the complex frequency solution.
-Consecutive cyclotron harmonics are increasingly unstable and are narrow due to the lack of Doppler shift associated with parallel drifts.
-(b) Lower panel: the same solutions as in panel (a) in the $(k_\parallel,\omega)$ plane where shading indicates growth rate.
+    indicates the parallel wavenumber in units of $\Omega_i/V_A$ of the complex
+    frequency solution. Consecutive cyclotron harmonics are increasingly
+    unstable and are narrow due to the lack of Doppler shift associated with
+    parallel drifts. (b) Lower panel: the same solutions as in panel (a) in the
+    $(\omega,k_\parallel)$ plane where shading indicates growth rate.
 The complex structure of the growth rates that exists in wavenumber space is projected into sharp harmonics on the frequency axis.}
@@
     parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions
-    in the $k_\parallel, \omega$ plane where shading indicates growth rate. Note
+    in the $(\omega,k_\parallel)$ plane where shading indicates growth rate. Note
     that the Doppler shift associated with the non-zero parallel drift speed
@@
 are the effect the spatial inhomogeniety of the magnetic field and background
-plasma, and the quantisation associatied with toroidal periodicity. These may
-create an eigenmode structure into which the MCI must fit.
+plasma, and the quantisation associatied with toroidal periodicity. These
+effects may create an eigenmode structure, not accounted for herein,
+associated with the kinetic MCI.
 Cottrell et al.
@@
 
-It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou useful discussions.
+It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou for useful discussions.
 

### 2022-02-06 manuscript.tex
updates post internal review

@@
 \begin{abstract}
-Ion cyclotron emission (ICE) originating from confined populations of fast ions in toroidal fusion plasmas is an important non-invasive, passive diagnostic for current and next generation devices.
-The ability to model the ICE signals accurately is an essential step towards inferring the characteristics of confined energetic alpha-particles or fast NBI ions and background plasma.
-In this paper, the linear growth rates of the magnetoacoustic cyclotron instability, which is the leading explanation for ICE, are calculated in high resolution 2D $(k_\parallel, k_\perp)$ space for parameters corresponding to JET Pulse No.
-26148.
-These calculations shed further light on the origin of doublet-splitting of the cyclotron harmonics as well as the apparent need to invoke nonlinear interactions with 1D3V particle-in-cell studies to account for lower harmonics with substantial amplitudes: the inclusion of signals of ICE from a well resolved range of propagation angles accounts for these effects.
+Ion cyclotron emission (ICE) originating from confined populations of fast ions
+  in toroidal fusion plasmas is an important non-invasive, passive diagnostic
+  for current and next generation devices. The ability to model the ICE signals
+  accurately is an essential step towards inferring the characteristics of
+  confined energetic alpha-particles or fast NBI ions and background plasma. In
+  this paper, the linear growth rates of the magnetoacoustic cyclotron
+  instability, which is the leading explanation for ICE, are calculated in high
+  resolution 2D $(k_\perp, k_\parallel)$ space for parameters corresponding to
+  JET Pulse No. 26148. These calculations shed further light on the origin of
+  doublet-splitting of the cyclotron harmonics as well as the apparent need to
+  invoke nonlinear interactions with 1D3V particle-in-cell studies to account
+  for lower harmonics with substantial amplitudes: the inclusion of signals of
+  ICE from a well resolved range of propagation angles can also account for
+  these effects.
 \end{abstract}
@@
 
-The foremost explanation of the mechanism responsible for ICE is the magnetoacoustic cyclotron instability (MCI)\cite{Belikov1976}, which arises from the free energy present in energetic ions with inverted distributions in velocity space.
-Theoretical understanding of ICE is based in analytical solutions to the linearised Maxwell-Vlasov system of equations\cite{Stix} where the energetic ion species was modelled as ring-beam for which $f(v_\parallel, v_\perp) \propto \exp(-(v_\parallel-u_\parallel)^2/ v_{th}^2)\delta(v_\perp - u_\bot)$ \cite{Dendy1994, McClements1996}.
-Here $u_\parallel$, $u_\perp$ and $v_{th}$ are the beam velocity (the signed drift speed along the magnetic field), ring speed (the speed perpendicular to the magnetic field) and thermal velocity, respectively.
-Subsequent studies augmented this understanding via nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one dimension and 3 vector components for velocities and electromagnetic field (i.e.
-1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}.
-These computational approaches also operate under the simplification that the plasma is both homogeneous and infinite albeit approximately via periodic boundary conditions.
-Linearised fluid approaches that incorporate toroidal geometry and inhomogeneous plasma and magnetic fields have shed light of the eigenmode structure of compressional Alfv{\'e}n eigenmodes, which like ICE are also fast Alfv{\'e}n modes driven by energetic ions, in spherical tokamaks\cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio tokamaks \cite{Kolesnichenko2000, Fulop2000, Heidbrink2006}.
+The foremost explanation of the mechanism responsible for ICE is the
+magnetoacoustic cyclotron instability (MCI)\cite{Belikov1976}, which arises from
+the free energy present in energetic ions with inverted distributions in
+velocity space. Theoretical understanding of ICE is based in analytical
+solutions to the linearised Maxwell-Vlasov system of equations\cite{Stix} where
+the energetic ion species was modelled as ring-beam for which $f(v_\parallel,
+v_\perp) \propto \exp(-(v_\parallel-u_\parallel)^2/ v_{th}^2)\delta(v_\perp -
+u_\bot)$ \cite{Dendy1994, McClements1996}. Here $u_\parallel$, $u_\perp$ and
+$v_{th}$ are the beam velocity (the signed drift speed along the magnetic
+field), ring speed (the speed perpendicular to the magnetic field) and thermal
+velocity, respectively. Subsequent studies augmented this understanding via
+nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and
+hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
+dimension and 3 vector components for velocities and electromagnetic field (i.e.
+1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These
+computational approaches also operate under the simplification that the plasma
+is both homogeneous and infinite albeit approximately via periodic boundary
+conditions. Linearised fluid approaches that incorporate toroidal geometry and
+inhomogeneous plasma and magnetic fields have shed light of the eigenmode
+structure of compressional Alfv{\'e}n eigenmodes, which like ICE are also fast
+Alfv{\'e}n modes driven by energetic ions, in spherical
+tokamaks\cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio
+tokamaks \cite{Kolesnichenko2000, Fulop2000, Heidbrink2006}.
 
-The effectiveness of analytical theory to reproduce the observed cyclotron harmonic peaks in the fluctuations of the magnetic field was first explained somewhat by 1D PIC simulations where the onset of nonlinearity appeared to imprint into the signal the amplitudes of each of the harmonics as they are at the end of the linear phase; each mode grows independently until growth is terminated for all modes when the fastest mode enters its nonlinear phase.
-Analysis shows that only higher cyclotron harmonics are destabilised\cite{McClements2015} to the MCI, however observations clearly indicate power at theoretically stable lower harmonics\cite{Cottrell1988}.
-In a further step towards accurate recreation of experimental observations, nonlinear phases of hybrid\cite{Carbajal2014} and fully kinetic\cite{Cook2013,Chapman2017} PIC simulations went on to show that wave-wave coupling was responsible for raising the amplitude of lower cyclotron harmonics.
-Most recently, hybrid 2D simulations have begun to shed light on the 2D3V dynamics of ICE\cite{Carbajal2021}.
+The effectiveness of analytical theory to reproduce the observed cyclotron
+harmonic peaks in the fluctuations of the magnetic field was first explained
+somewhat by 1D PIC simulations. In these simulations the onset of nonlinearity
+appeared to imprint the amplitudes of each of the harmonics into the signal as
+they are at the end of the linear phase; each mode grows independently until 
+growth is terminated for all modes when the fastest mode enters its nonlinear
+phase meaning that both the amount of free energy available for instability has
+decreased and that the unstable energy ion distribution function has been
+perturbed into a more stable state. Analysis shows that only higher cyclotron
+harmonics are destabilised\cite{McClements2015} to the MCI, however observations
+clearly indicate power at theoretically stable lower
+harmonics\cite{Cottrell1988}. In a further step towards accurate recreation of
+experimental observations, nonlinear phases of hybrid\cite{Carbajal2014} and
+fully kinetic\cite{Cook2013,Chapman2017} PIC simulations went on to show that
+wave-wave coupling was responsible for raising the amplitude of lower cyclotron
+harmonics. Most recently, hybrid 2D simulations have begun to shed light on the
+2D3V dynamics of ICE\cite{Carbajal2021}.
 
-To better enable ICE as a diagnostic for confined fast ions, it is necessary to further develop our understanding with a range of tools.
-In this work, I present the 2D $(k_\parallel, k_\perp)$ growth rates of the MCI, under the conditions of the JET PTE\cite{Cottrell1993} namely Pulse No.
-26148 are presented.
-These growth rates are obtrained from a dispersion relation solver, PlasmaDispersionRelations.jl, written in the Julia programming language\cite{Bezanson2017}.
-These results narrow the gap between analytical theory and expensive HPC simulations and can be run readily on desktop computers in a matter of minutes, further enabling research into this topic.
+To better enable ICE as a diagnostic for confined fast ions, it is necessary to
+further develop our understanding with a range of tools. In this work, I present
+the 2D $(k_\perp, k_\parallel)$ growth rates of the MCI under the conditions of
+outer mid-plane of the JET PTE\cite{Cottrell1993} Pulse No. 26148. These growth
+rates are obtained from a dispersion relation solver,
+PlasmaDispersionRelations.jl, written in the Julia programming
+language\cite{Bezanson2017}. These results narrow the gap between analytical
+theory and expensive HPC simulations and can be run readily on desktop computers
+in a matter of minutes, further enabling research into this topic. Note that
+this study focusses on edge-ICE, named for the region from which the signal
+arises, as opposed to core-ICE\cite{Ochoukov2018}.
 
@@
 deuteron ions. Calculations are made for alpha-particles with ring-beams of two
-pitch-angle cosine values: $u_\parallel=-0.646 u_0 \approx
--\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_1})) \simeq -1.1 V_A$, which is
+values of pitch-angle cosine: $u_\parallel=-0.646 u_0 \approx
+-\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_{ICE}})) \simeq -1.1 V_A$, which is
 associated with the ion-cyclotron emission observed on JET Pulse No. 26148 for
-which $R_0=3.46m$ and
-$\psi_0=55.2^\circ$ are the major radius and angle pitch angle at birth, and
-$R_1=4m$ is the major radius of the emission region, see
+which $R_0=3.46$ m and
+$\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
+$R_{ICE}=4$ m is the major radius of the emission region, see
 Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
 $u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
-Fig.
-1(a) of Ref \cite{Cook2013}.
-Here $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is the speed of an alpha-particle with an energy of 3.6 MeV.
-The complex frequency solutions, $\omega$, are found corresponding to the fast Alfv{\'e}n branch on a grid of $(k_\parallel, k_\bot)$.
+Fig. 1(a) of Ref \cite{Cook2013}. Here, $V_A$ is the Alfv{\'en} speed and $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is
+the speed of an alpha-particle with an energy of 3.6 MeV. The complex frequency
+solutions, $\omega$, along the fast Alfv{\'e}n branch are found on a grid of $(k_\parallel, k_\bot)$.
 
@@
 of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
-signal of Fig. 2 (ibid). The electrons and deuterons, both modelled as
-Maxwellians, have temperatures of 1 keV and the electrons have a number density
+signal of Fig. 2 (ibid). The electrons and deuterons have temperatures of 1 keV and the electrons have a number density
 of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
-$\mathrm{T}$, which corresponds a major radius of $4m$.
-The ring-beam distribution function, Eq.
-\ref{eq:alphaf}, is a separable function of $v_\parallel$, i.e.
-the beam given by Eq.
-1, and $v_\perp$, i.e.
-the ring given by Eq.
-2.
-The component parallel (perpendicular) to the magnetic field is a drifting Maxwellian with thermal speed $v_{\parallel, th}$ ($v_{\perp, th}$) and drift velocity $u_\parallel$ ($u_\perp$).
-The thermal spread of the alpha particles' distribution function $v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$.
-The concentration ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.
+$\mathrm{T}$, which corresponds a major radius of $4$ m.
+The ring-beam distribution function, Eq. \ref{eq:alphaf}, is a separable
+function of the beam in $v_\parallel$ given by Eq. 1 and the ring in $v_\perp$
+given by Eq. 2. This distribution function is parameterised by the drift 
+velocity parallel to the magnetic field $u_\parallel$, the drift speed 
+perpendicular to the magnetic
+field $u_{\perp}$, and the thermal velocity in both directions $v_{th}$. The
+thermal spread of the alpha particles' distribution function 
+$v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$. The number density
+ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.
 
@@
 
-Solutions to the linearised Maxwell-Vlasov system of equations are given by the dispersion relation, Eq.
-\ref{eq:dettensor}, which is satisfied for sets of $(\omega,\vec{k})$.
-The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$,  given by Eq.
-\ref{eq:dielectrictensor} is calculated for each species.
+Solutions to the linearised Maxwell-Vlasov system of equations are given by the
+dispersion relation, Eq. \ref{eq:dettensor}, which is satisfied for sets of
+$(\omega,\vec{k})$ given the distribution functions of all the species present.
+The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$, Eq. \ref{eq:dielectrictensor} is calculated for each species.
 
@@
 
-\noindent $J_n$ is the $n$th Bessel function of the first kind and takes the argument $z_s=\frac{k_\perp v_\perp}{\Omega_s}$ where $J_n'=\frac{\partial J_n}{\partial z}$.
+\noindent $J_n$ is the $n$th Bessel function of the first kind taking the
+argument $z_s=k_\perp v_\perp / \Omega_s$ and $J_n'=\frac{\partial J_n}{\partial z}$.
 
-The integrals over perpendicular velocity required for calculating the dielectric tensor for Maxwellian distribution functions are given in terms of modified Bessel functions of the first kind for both deuterons and ions i.e.
-when $u_\perp =0$.
-The same perpendicular integrals over the alpha-particles' ring distribution is calculated numerically via Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK} library.
-An absolute and relative tolerance of $0$ and $10^{-8}$, respectively, on the value of the integrals is used.
+The integrals over perpendicular velocity required for calculating the
+dielectric tensor for Maxwellian distribution functions are given in terms of
+modified Bessel functions of the first kind\cite{Stix} for both electrons and
+ldeuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
+alpha-particles' ring distribution is calculated numerically via
+Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK}
+library. An absolute and relative tolerance of $0$ and $10^{-8}$, respectively,
+on the value of the integrals is used.
 
-Since all the parallel distribution functions for all species are Maxwellians, the parallel integrals can be calculated from the well known plasma dispersion function\cite{Fried1961}, which is given in the appendix.
+Since all the parallel distribution functions for all species are Maxwellians,
+the parallel integrals can be calculated from the well known plasma dispersion
+function\cite{Fried1961}, which is laid out in the appendix.
 
-The summation over Bessel function indices continues until the relative change of the L2 norm between the current and previous iteration is less than a relative tolerance of $10^{-8}$.
-Root finding of Eq.
-\ref{eq:dettensor} is performed by an amalgamation of the Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method, whereby if the Nelder-Mead simplex does not contain a root, as calculated via the winding method, then the optimisation procedure follows the Nelder-Mead method, otherwise if a root is present with the simplex then the simplex bifurcates and the child simplex that contains the root is used for the next iteration.
-The optimisation iteration procedure finishes when the real extent of the simplex is less than $10^{-4}$ of the alpha-particle cyclotron frequency $\Omega_i$ (the only ion species considered are deuterons and alpha-particles, which have identical cyclotron frequencies) and the imaginary extent is less than $10^{-5}\,\Omega_i$.
-This approach is found to be adequately robust and faster than the Nelder-Mead method or winding number approach used in isolation of one another.
-Each 512 by 512 scan in $(k_\parallel, k_\perp)$ takes between 5 and 6 minutes on a laptop.
+The summation over Bessel function indices in Eq. \ref{eq:dielectrictensor}
+continues until the relative change of the L2 norm between the current and
+previous iteration is less than a tolerance of $10^{-8}$. Root finding of Eq.
+\ref{eq:dettensor} is performed by an amalgamation of the
+Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method,
+whereby if the Nelder-Mead simplex does not contain a root, as calculated via
+the winding method, then the optimisation procedure follows the Nelder-Mead
+method, otherwise a non-zero winding number indicates that a root is present
+within the simplex and the simplex
+bifurcates. The child simplex that contains the root is used for the next
+iteration. The optimisation iteration procedure finishes when the real extent of
+the simplex is less than $10^{-4}\, \Omega_i$ and the imaginary extent is less
+than $10^{-5}\,\Omega_i$, where $\Omega_i$ is the cyclotron frequency of both
+the alpha-particles and deuterons since they share a charge-mass ratio.
+This approach is found to be adequately robust and
+faster than the Nelder-Mead method or winding number approach used in isolation
+of one another. Each 512 by 512 scan in $(k_\perp, k_\parallel)$ takes between 5
+and 6 minutes on a laptop.
 
@@
 rates. The simplicity of the consecutive unstable cyclotron harmonics visible in
-panel (a) betrays the complexity of the structure in the $(\omega,k_\parallel)$ plane.
-Integration over $k_\parallel$ projects the solutions onto the frequency axis, which loses the structure and produces narrow harmonics.
-Fig.
-\ref{fig:2D_imag_zeropitch} plots shows the same data as Fig.
-\ref{fig:2D_F12_zeropitch} with damped and normal modes in addition, except shading indicates the growth rates in $(k_\parallel, k_\perp)$ space.
-Overlaid contours indicate the real frequency in units of ion cyclotron harmonics.
-Only narrow regions around the cyclotron harmonics are unstable.
-Overlaid straight traces indicate propagation angle.
-No Doppler shift is present due to the lack of parallel drift in the alpha-particle ring-beam.
-The growth rate increases approximately linearly with cyclotron harmonic number, which is consistent with the literature\cite{Dendy1994}.
-Together these figures serves to introduce a 2D view (growth rates in $(k_\parallel, k_\perp)$ space) of traditionally 1D data (growth rates against frequency).
-This will become important when a non-zero $u_\parallel$ is taken into account next.
+panel (a) betrays the complexity of the structure in the $(\omega,k_\parallel)$
+plane. Integration over $k_\parallel$ projects the solutions onto the frequency
+axis, which loses the structure and produces narrow harmonics. Figure
+\ref{fig:2D_imag_zeropitch} shows the same data as Fig.
+\ref{fig:2D_F12_zeropitch} in $(k_\perp, k_\parallel)$ space with damped and
+normal modes in addition, except shading indicates the growth rates. Overlaid
+contours indicate the real frequency in units of ion cyclotron harmonics. Only
+narrow regions around the cyclotron harmonics are unstable. Overlaid straight
+traces indicate propagation angle. No Doppler shift is present due to the lack
+of parallel drift in the alpha-particle ring-beam. The growth rate increases
+approximately linearly with cyclotron harmonic number, which is consistent with
+the literature\cite{Dendy1994}. Together these figures serve to introduce a 2D
+view (growth rates in $(k_\perp, k_\parallel)$ space) of traditionally 1D data
+(growth rates against frequency). This will become important when a non-zero
+$u_\parallel$ is taken into account next.
 
@@
     alpha-particles i.e. when there is zero parallel drift, $u_\parallel=0$ in
-    Eq. 1. (a) Upper panel: growth rate as a function of frequency where shading
-    indicates the parallel wavenumber in units of $\Omega_i/V_A$ of the complex
-    frequency solution. Consecutive cyclotron harmonics are increasingly
-    unstable and are narrow due to the lack of Doppler shift associated with
-    parallel drifts. (b) Lower panel: the same solutions as in panel (a) in the
-    $(\omega,k_\parallel)$ plane where shading indicates growth rate.
-The complex structure of the growth rates that exists in wavenumber space is projected into sharp harmonics on the frequency axis.}
+    Eq. 1. (a) Upper panel: growth rate as a function of real frequency where shading
+    indicates the parallel wavenumber, $k_\parallel$, in units of $\Omega_i/V_A$
+    of the complex frequency solution. Consecutive cyclotron harmonics are
+    increasingly unstable and are narrow due to the lack of Doppler shift 
+    associated with parallel drifts. (b) Lower panel: the same solutions
+    as in panel (a) in the $(\omega,k_\parallel)$ plane where shading indicates 
+    growth rate. The complex structure of the growth rates that exists in
+    wavenumber space is projected into sharp harmonics on the frequency axis.}
     \label{fig:2D_F12_zeropitch}
@@
 \includegraphics[scale=0.9]{fig2.pdf}
-    \caption{(colour online) Shading indicates the growth rate of fast Alfv{\'e}n waves excited by a ring of fusion born alpha-particles, as a function of parallel wavenumber on the ordinate and perpendicular wavenumber on the abscissa.
-No solution is found where shading is absent.
-The alpha-particle distribution function is given by Eq.
-\ref{eq:alphaf} with $u_{\parallel} = 0$.
-Consecutive straight traces indicate the propagation angle with respect to the magnetic field: the horizontal line is exactly perpendicular; solid traces fan out at intervals of five degrees; and dashed traces indicate intervals of one degree.
-Curved grey traces indicate the location in $(k_\parallel, k_\perp)$ of the cyclotron harmonics (i.e.
-the real component of the complex $\omega$ solutions), which are annotated along the left hand side and upper edge of the figure.
-Frequencies are in units of the alpha-particle cyclotron frequency, and wavenumbers are in alpha-particle cyclotron frequencies per Alfv{\'e}n speed.
-Note the up-down symmetry due to the lack of parallel drift in the energetic alpha particles.}
+    \caption{(colour online) Shading indicates the growth rate of fast
+    Alfv{\'e}n waves excited by a ring of fusion born alpha-particles, as a
+    function of parallel wavenumber on the ordinate and perpendicular wavenumber
+    on the abscissa. No solution is found where shading is absent. The
+    alpha-particle distribution function is given by Eq. \ref{eq:alphaf} with
+    $u_{\parallel} = 0$. Consecutive straight traces indicate the propagation
+    angle with respect to the magnetic field: the horizontal line is exactly
+    perpendicular; solid traces fan out at intervals of five degrees; and dashed
+    traces indicate intervals of one degree. Curved grey traces indicate the
+    location in $(k_\perp, k_\parallel)$ of the cyclotron harmonics (i.e. the
+    real component of the complex $\omega$ solutions), which are annotated along
+    the left hand side and upper edge of the figure. Frequencies are in units of
+    the alpha-particle cyclotron frequency, and wavenumbers are in
+    alpha-particle cyclotron frequencies per Alfv{\'e}n speed. Note the up-down
+    symmetry due to the lack of parallel drift in the energetic alpha
+    particles.}
     \label{fig:2D_imag_zeropitch}
@@
 
-Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physical pitch-angle
+Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physically more
+realistic pitch-angle
 case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
@@
 strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
-of the $11^{th}$ harmonic is observed: see Fig. 2 of Cottrell 1993 for a single
+of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
 signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
@@
 wavenumber space. The following describes how these three features are
-connected. The doublet splitting arises due to the dip in growth rate where
+connected. Doublet splitting arises due to the dip in growth rate along the 
+band of instability where
 $k_\parallel$ goes through zero combined with the change in sign of the Doppler
 shift of the real frequency as $k_\parallel$ changes sign. The effect is
-exaggerated for higher harmonics where the growth rate diminishes to zero for
+larger for higher harmonics where the growth rate diminishes to zero for
 perpendicular propagation; see e.g. the $12^{th}$ harmonic in Fig.
 \ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
-k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $\Omega_{i}$ is the ion cyclotron
-frequency and $V_A$ is the Alfv{\'e}n speed. Other possible mechanisms for
+k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed. Other possible mechanisms for
 doublet splitting could be grad-B or curvature
-drifts\cite{Cottrell1993,Fulop1998}.
-Qualitatively, Fig.
-\ref{fig:2D_F12_physicalpitch}(a) resembles Fig.
-2 of Cottrell 1993 in that the signal below $\sim 7$ cyclotron harmonics is lower and flatter with well defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change larger and the harmonics are less easily identifiable.
-These features can only be recreated with a 2D calculation for these distributions functions.
-Note that Fig.
-\ref{fig:2D_imag_physicalpitch} shows these two different regimes more clearly.
-The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is characterised by narrow regions of instability near Doppler shifted cyclotron harmonics that are stabilised when $k_\parallel=0$ but unstable at larger negative parallel wavenumbers.
-The higher frequency behaviour in the region $\omega\geq 7\Omega_i$ is dominated by broader regions of instability in frequency and wavenumber near to Doppler shifted cyclotron harmonics, which have a complex structure in wavenumber space.
-Modes propagating exactly perpendicular become stable in the region $k_\perp > 13 \Omega_i/V_A$ whilst obliquely propagating modes remain unstable; see Fig.
-9 of Dendy et al 1995, which plots the analytic expression for the growth rate of the magnetoacoustic cylcotron instability, i.e.
-the instability thought responsible for ICE, at 5 slices of propagation angle.
-The Doppler shift of the growing modes can reach $\approx \Omega_i$ for large propagation angles at lower frequencies and also for moderately oblique angles at higher frequencies; see, for example, the mode associated with the $15^{th}$ cyclotron harmonic at, $(k_\parallel \simeq -\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e.
-a propagation angle of $93^\circ$).
-Furthermore, the breadth of the harmonics can be attributed to the contributions from multiple propagation angles even though the lower harmonics are stable for forward travelling and perpendicular propagation angles.
+drifts\cite{Cottrell1993,Fulop1998}. Qualitatively, Fig.
+\ref{fig:2D_F12_physicalpitch}(a) resembles Fig. 2 of Cottrell 1993 in that the
+signal below the $7^{th}$ cyclotron harmonics is lower and flatter with well
+defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change
+larger and the harmonics are less easily identifiable. These features cannot be
+be recreated with a 1D calculation for these distributions functions. Note that
+Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes more
+clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is
+characterised by narrow regions of instability near Doppler shifted cyclotron
+harmonics that are stabilised when $k_\parallel=0$ but unstable at larger
+negative parallel wavenumbers. The higher frequency behaviour in the region
+$\omega\geq 7\Omega_i$ is dominated by broader regions of instability in
+frequency and wavenumber near to Doppler shifted cyclotron harmonics, which have
+a complex structure in wavenumber space. Modes numbering 1, 2, 6, 7, 8, 12, and
+13 are stable when propagating exactly perpendicular whilst obliquely
+propagating modes remain unstable; see Fig. 9 of Dendy et al 1995, which plots
+the analytic expression for the growth rate of the MCI at 5 slices of
+propagation angle. The Doppler shift of the growing modes can reach $\approx
+\Omega_i$ for large propagation angles at lower frequencies and also for
+oblique angles at higher frequencies; see, for example, the mode
+associated with the $15^{th}$ cyclotron harmonic near $k_\parallel \simeq
+-\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e. a propagation
+angle of $93^\circ$. Furthermore, the breadth of the harmonics can be
+attributed to the contributions from multiple propagation angles even though the
+lower harmonics are largely stable for forward travelling and perpendicular
+propagation angles.
 
@@
     \caption{(colour online) The imaginary component of the frequencies for the
-    case with parallel drift, $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
+    case with parallel drift $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
     Panel (a) plots growth rate against frequency where shading indicates
@@
     broadens the unstable peaks in frequency. The $10^{th}$ and $11^{th}$ cyclotron
-    harmonics are split into doublets, i.e. they have two peaks because both
-    the growth rate and real frequency vary with wavenumber. Compare with
+    harmonics are split into doublets, i.e. they have two peaks. Compare with
     Fig. 2 of Cottrel et al 1993.}
@@
 \includegraphics[scale=0.9]{fig4.pdf}
-    \caption{(colour online) As in Fig.
-\ref{fig:2D_imag_zeropitch}, except for a ring-beam of alpha particles, with a physical pitch-angle value resulting in $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
-The non-zero parallel drift speed causes up-down asymmetry; this figure for a pitch-angle cosine of the opposite sign, $u_{\parallel} = 0.646 u_0$, is identical expect flipped about $k_\parallel=0$.}
+    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except for
+    a ring-beam of alpha particles with a more realistic pitch-angle value
+    $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$. The non-zero parallel drift
+    speed causes up-down asymmetry; this figure for a pitch-angle cosine of the
+    opposite sign, $u_{\parallel} = 0.646 u_0$, is identical expect flipped
+    about $k_\parallel=0$.}
     \label{fig:2D_imag_physicalpitch}
@@
 \includegraphics[scale=0.9]{fig5.pdf}
-    \caption{(colour online) As in Fig.
-\ref{fig:2D_imag_zeropitch}, except that the shading indicates the real part of the smoothed solution normalised by the high frequency expression for the fast Alfv{\'e}n wave set out by Eq.
-25 of Ref.
-\cite{Dendy1994}, which is also Eq.
-11 in Ref.
-\cite{McClements1996}.
-The differences in real frequency for the two sets of calculations, physical pitch-angle cosine of $-0.646$ and zero, are negligible.}
+    \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except that
+    the shading indicates the real part of the smoothed solution normalised by
+    the high frequency expression for the fast Alfv{\'e}n wave set out in Eq. 25
+    of Ref. \cite{Dendy1994}, which is also Eq. 11 in Ref.
+    \cite{McClements1996}. The differences in real frequency for the two sets of
+    calculations, a pitch-angle cosine of $-0.646$ and zero, are
+    negligible.}
     \label{fig:2D_real_zeropitch}
@@
 
-This paper presents the linear Maxwell-Vlasov dispersion relation of a plasma consisting of thermal electrons, thermal deuteron fuel ions and a minority energetic ring-beam of fusion born alpha-particles.
-These high resolution $(k_\parallel, k_\perp)$ calculations revisit ICE observations from JET Pulse No.
-26148\cite{Cottrell1993,Dendy1995} and reproduce several important features: the step change in signal strength from $\omega < 6$ to $\omega > 6$; the doublet splitting of peaks at $\omega \approx 10\Omega_i$ on Fig.
-\ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\parallel, k_\perp)$ structure in Fig.
-\ref{fig:2D_imag_physicalpitch}.
-The Doppler shift for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron frequency, which makes it difficult to determine which parts of the signal come from which cyclotron harmonic numbers.
-Further effort is required to account for the extra doublets visible in the continuous trace associated with JET Pulse No.
-26148 in Fig.
-2 of Ref.
-\cite{Cottrell1993}.
+This paper presents the linear Maxwell-Vlasov dispersion relation of a plasma
+consisting of thermal electrons, thermal deuteron fuel ions and a minority
+energetic ring-beam of fusion born alpha-particles. These high resolution
+$(k_\perp, k_\parallel)$ calculations revisit ICE observations from JET Pulse
+No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
+the step change in signal strength from $\omega < 6$ to $\omega > 6$; the
+doublet splitting of peaks at $\omega \approx 10\Omega_i$ in Fig.
+\ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\parallel,
+k_\perp)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift
+for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron
+frequency, which makes it difficult to determine which parts of the signal come
+from which cyclotron harmonic numbers. Further effort is required to account for
+the extra doublets visible in the continuous trace associated with JET Pulse No.
+26148 in Fig. 2 of Ref. \cite{Cottrell1993}.
 
@@
 like those presented here, are faster. It has been shown that it is possible to
-recreate signals containing lower harmonics of substantial amplitudes by
+qualitatively recreate signals containing lower harmonics of substantial amplitudes by
 invoking 2D linear physics. The presence of ICE signals at lower harmonics in
@@
 There are several important features missing from these calculations. Among them
-are the effect the spatial inhomogeniety of the magnetic field and background
+are the effects of the spatial inhomogeniety of the magnetic field and background
 plasma, and the quantisation associatied with toroidal periodicity. These
@@
 associated with the kinetic MCI.
-Cottrell et al.
-\cite{Cottrell1993} does discuss the implications of propagation angles on ICE signals and indicate that lower harmonics must propagate increasingly close to exactly perpendicular.
-These effects remain the subject of further work to incorporate them in linear and nonlinear studies of ICE.
+Cottrell et al. \cite{Cottrell1993} does discuss the implications of propagation
+angles on ICE signals and indicate that lower harmonics must propagate
+increasingly close to exactly perpendicular. These effects remain the subject of
+further work to incorporate them in linear and nonlinear studies of ICE.
 
-This work showcases a new high $(\omega,k)$ resolution tool for studying ICE and captures the fine detail of ICE signals, which is challenging to capture with nonlinear PIC code simulations since they suffer from low resolution in frequency and wavenumber space due to the phenomenon's transience and computational expense, respectively.
-As such, this work represents a complementary tool that builds upon purely analytical work and nonlinear PIC studies, which in turn enhances the feasibility of exploiting ICE as a passive diagnostic of confined and leaving ions in magnetised fusion plasmas.
+This work showcases a new high resolution tool for studying ICE and captures the
+fine detail of ICE signals, which is challenging to capture with nonlinear PIC
+code simulations since they suffer from low resolution in frequency and
+wavenumber space due to the phenomenon's transience and computational expense,
+respectively. As such, this work represents a complementary tool that builds
+upon purely analytical work and nonlinear PIC studies, which in turn enhances
+the feasibility of exploiting ICE as a passive diagnostic of confined and
+leaving ions in magnetised fusion plasmas.
 

### 2022-02-16 manuscript.tex
these changes got lost before swapping from overleaf

@@
 hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
-dimension and 3 vector components for velocities and electromagnetic field (i.e.
-1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These
-computational approaches also operate under the simplification that the plasma
-is both homogeneous and infinite albeit approximately via periodic boundary
-conditions. Linearised fluid approaches that incorporate toroidal geometry and
+spatial dimension and three vector components for velocities and electromagnetic
+field (i.e. 1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These
+computational approaches operate under similar simplifications as the analytical
+treatments in that the plasma is initially spatially homogeneous and periodic
+in the spatial domain.
+Linearised fluid approaches that incorporate toroidal geometry and
 inhomogeneous plasma and magnetic fields have shed light of the eigenmode
@@
 perturbed into a more stable state. Analysis shows that only higher cyclotron
-harmonics are destabilised\cite{McClements2015} to the MCI, however observations
-clearly indicate power at theoretically stable lower
+harmonics are destabilised\cite{McClements2015} to the MCI, however experimental
+observations clearly indicate power at theoretically stable lower
 harmonics\cite{Cottrell1988}. In a further step towards accurate recreation of
@@
 further develop our understanding with a range of tools. In this work, I present
-the 2D $(k_\perp, k_\parallel)$ growth rates of the MCI under the conditions of
+the growth rates in 2D wavenumber space of the MCI under the conditions of
 outer mid-plane of the JET PTE\cite{Cottrell1993} Pulse No. 26148. These growth
@@
 in a matter of minutes, further enabling research into this topic. Note that
-this study focusses on edge-ICE, named for the region from which the signal
+this study focusses on edge-ICE, named for the region of plasma from which the signal
 arises, as opposed to core-ICE\cite{Ochoukov2018}.
@@
 Solutions to the linearised Maxwell-Vlasov system of equations are found in
-$(k_\parallel, k_\bot)$ space for a ring-beam distribution of energetic minority
+parallel and perpendicular wavenumber space, $(k_\parallel, k_\bot)$, for a
+ring-beam distribution of energetic minority
 alpha-particles and Maxwellian electrons and background thermal majority
@@
 of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
-signal of Fig. 2 (ibid). The electrons and deuterons have temperatures of 1 keV and the electrons have a number density
+signal of Fig. 2 in the same reference. The electrons and deuterons have
+temperatures of 1 keV and the electrons have a number density
 of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
 $\mathrm{T}$, which corresponds a major radius of $4$ m.
-The ring-beam distribution function, Eq. \ref{eq:alphaf}, is a separable
-function of the beam in $v_\parallel$ given by Eq. 1 and the ring in $v_\perp$
-given by Eq. 2. This distribution function is parameterised by the drift 
+The ring-beam distribution function, Eq. 1, is a separable
+function of the beam in $v_\parallel$ given by Eq. 2 and the
+ring in $v_\perp$ given by Eq. 3.
+This distribution function is parameterised by the drift
 velocity parallel to the magnetic field $u_\parallel$, the drift speed 
@@
 \begin{align}
-f_{\alpha,\parallel}\left(v_\parallel\right) &= \frac{\exp\left(-\frac{(v_\parallel - u_\parallel)^2}{v_{\parallel,th}^2}\right)}{v_{\parallel,th} \sqrt{2 \pi}}\\
-f_{\alpha,\perp}\left(v_\perp\right) &= \frac{1}{2\pi}\frac{\exp\left(-\frac{\left(v_\bot - u_\bot\right)^2}{v_{\bot,th}^2}\right)}{\left(\frac{\sqrt \pi}{2} v_{\perp, th} u_\bot (1 - \mathrm{erf}(-\frac{u_\bot}{v_{\perp,th}})) + \exp(-u_\bot^2 / v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}\\
-f_\alpha\left(v_\parallel, v_\perp\right) &= f_{\alpha,\parallel}\left(v_\parallel\right)f_{\alpha,\perp}\left(v_\perp\right)
-    \label{eq:alphaf}
+f_\alpha\left(v_\parallel, v_\perp\right) &=
+  f_{\alpha,\parallel}\left(v_\parallel\right)f_{\alpha,\perp}\left(v_\perp\right)\\
+f_{\alpha,\parallel}\left(v_\parallel\right) &=
+  \frac{\exp\left(-\frac{(v_\parallel -
+  u_\parallel)^2}{v_{\parallel,th}^2}\right)}{v_{\parallel,th} \sqrt{2
+  \pi}}\\
+f_{\alpha,\perp}\left(v_\perp\right) &=
+  \frac{1}{2\pi}\frac{\exp\left(-\frac{\left(v_\bot -
+  u_\bot\right)^2}{v_{\bot,th}^2}\right)}{\left(\frac{\sqrt \pi}{2} v_{\perp,
+  th} u_\bot (1 - \mathrm{erf}(-\frac{u_\bot}{v_{\perp,th}})) + \exp(-u_\bot^2 /
+  v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}\\
 \end{align}
@@
 The summation over Bessel function indices in Eq. \ref{eq:dielectrictensor}
-continues until the relative change of the L2 norm between the current and
+continues until the relative change between the current and
 previous iteration is less than a tolerance of $10^{-8}$. Root finding of Eq.
-\ref{eq:dettensor} is performed by an amalgamation of the
+\ref{eq:dettensor} in complex frequency plane is performed by an amalgamation of the
 Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method,
-whereby if the Nelder-Mead simplex does not contain a root, as calculated via
-the winding method, then the optimisation procedure follows the Nelder-Mead
-method, otherwise a non-zero winding number indicates that a root is present
-within the simplex and the simplex
-bifurcates. The child simplex that contains the root is used for the next
+whereby if the values at the nodes of the Nelder-Mead simplex, which in this
+case is just a triangle, have a winding number
+of zero and therefore the simplex does not contain a root then the optimisation
+procedure follows the Nelder-Mead method, otherwise the simplex bifurcates.
+The child simplex that contains the root is used for the next
 iteration. The optimisation iteration procedure finishes when the real extent of
@@
 axis, which loses the structure and produces narrow harmonics. Figure
-\ref{fig:2D_imag_zeropitch} shows the same data as Fig.
-\ref{fig:2D_F12_zeropitch} in $(k_\perp, k_\parallel)$ space with damped and
-normal modes in addition, except shading indicates the growth rates. Overlaid
+\ref{fig:2D_imag_zeropitch} gives $(k_\perp, k_\parallel)$ space view of the
+same data as Fig. \ref{fig:2D_F12_zeropitch} as well as damped and
+normal modes, except shading indicates the growth rates. Overlaid
 contours indicate the real frequency in units of ion cyclotron harmonics. Only
@@
 view (growth rates in $(k_\perp, k_\parallel)$ space) of traditionally 1D data
-(growth rates against frequency). This will become important when a non-zero
-$u_\parallel$ is taken into account next.
+(growth rates against frequency). The importance of a non-zero $u_\parallel$ is
+discussed next.
 
@@
     on the abscissa. No solution is found where shading is absent. The
-    alpha-particle distribution function is given by Eq. \ref{eq:alphaf} with
+    alpha-particle distribution function is given by Eq. 1 with
     $u_{\parallel} = 0$. Consecutive straight traces indicate the propagation
@@
 case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
-\ref{fig:2D_F12_zeropitch}. There are several points worth noting. First, the
+\ref{fig:2D_F12_zeropitch} except there is an additional black trace in the
+upper panel, which represents the measured JET ICE intensity signal; the continuous
+trace from Fig. 2 in Ref. \cite{Cottrell1993}.
+There are several points worth noting. First, the
 growth rate is no longer linear with cyclotron harmonic number. Second,
@@
 larger for higher harmonics where the growth rate diminishes to zero for
-perpendicular propagation; see e.g. the $12^{th}$ harmonic in Fig.
-\ref{fig:2D_imag_physicalpitch} near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
-k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed. Other possible mechanisms for
-doublet splitting could be grad-B or curvature
-drifts\cite{Cottrell1993,Fulop1998}. Qualitatively, Fig.
-\ref{fig:2D_F12_physicalpitch}(a) resembles Fig. 2 of Cottrell 1993 in that the
+perpendicular propagation.
+
+This effect can be seen in Fig. \ref{fig:2D_imag_physicalpitch}, which is the
+counterpart figure to Fig. \ref{fig:2D_F12_physicalpitch} as Fig.
+\ref{fig:2D_imag_zeropitch} is to Fig. \ref{fig:2D_F12_zeropitch},
+at e.g. the $12^{th}$ harmonic  near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
+k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed.
+Other possible mechanisms for doublet splitting could be grad-B or curvature
+drifts\cite{Cottrell1993,Fulop1998}. Qualitatively, scattered points in Fig.
+\ref{fig:2D_F12_physicalpitch}(a) resembles the experimental observations given
+by the black trace in that the
 signal below the $7^{th}$ cyclotron harmonics is lower and flatter with well
@@
 characterised by narrow regions of instability near Doppler shifted cyclotron
-harmonics that are stabilised when $k_\parallel=0$ but unstable at larger
+harmonics that are stable when $k_\parallel=0$ and unstable at larger
 negative parallel wavenumbers. The higher frequency behaviour in the region
@@
 13 are stable when propagating exactly perpendicular whilst obliquely
-propagating modes remain unstable; see Fig. 9 of Dendy et al 1995, which plots
-the analytic expression for the growth rate of the MCI at 5 slices of
-propagation angle. The Doppler shift of the growing modes can reach $\approx
-\Omega_i$ for large propagation angles at lower frequencies and also for
+propagating modes remain unstable. The magnitude of the Doppler shift of
+the growing modes can reach $\approx \Omega_i$ for large propagation angles at 
+lower frequencies and also for
 oblique angles at higher frequencies; see, for example, the mode
@@
     Panel (a) plots growth rate against frequency where shading indicates
-    parallel wavenumber in units of $\Omega_i/V_A$, and (b) plots the solutions
+    parallel wavenumber in units of $\Omega_i/V_A$ and the black trace shows
+    the re-scaled ICE signal from JET Pulse \#21648 digitised from Fig. 2 of
+    Cottrell et al 1993 (with permission from IOP Publishing), (b) plots the solutions
     in the $(\omega,k_\parallel)$ plane where shading indicates growth rate. Note
@@
     broadens the unstable peaks in frequency. The $10^{th}$ and $11^{th}$ cyclotron
-    harmonics are split into doublets, i.e. they have two peaks. Compare with
-    Fig. 2 of Cottrel et al 1993.}
+    harmonics are split into doublets, i.e. they have two peaks.}
     \label{fig:2D_F12_physicalpitch}
@@
     speed causes up-down asymmetry; this figure for a pitch-angle cosine of the
-    opposite sign, $u_{\parallel} = 0.646 u_0$, is identical expect flipped
+    opposite sign, $u_{\parallel} = 0.646 u_0$, is identical except flipped
     about $k_\parallel=0$.}
@@
 more powerful higher harmonics\cite{Carbajal2014}. This work suggests that
-observations of lower ICE harmonics may arise either due to the effects of 2D
-linear physics or to nonlinear 1D physics\cite{Chapman2018}, or a combination.
+observations of lower ICE harmonics may arise due to 2D
+linear physics effects but does not exclude the possiblity that nonlinear 1D
+physics\cite{Chapman2018} may also play a part.
 There are several important features missing from these calculations. Among them

### 2022-02-16 manuscript.tex
justify comparing observations in dB and linear growth rate

@@
 trace from Fig. 2 in Ref. \cite{Cottrell1993}.
-There are several points worth noting. First, the
+The observations of ICE intensity are plotted on a log scale scaled to match
+the range of the linear MCI growth rates. This juxtaposition is reasonable since
+nonlinear PIC code simulations indicate that all growing modes stop when the
+fastest growing mode becomes nonlinear at time $\tau$. Under the assumption
+that all modes start growing in unison from identical seed amplitudes then the
+$log$ of the amplitude of a mode with growth rate $\gamma$ is proportional to
+$\gamma \tau$, allowing for a comparison with observations also on a log scale.
+There are several points worth noting about this figure. First, the
 growth rate is no longer linear with cyclotron harmonic number. Second,

### 2022-02-26 manuscript.tex
title change, refs names fixup

@@
 
-\title[]{Doublet splitting and linear amplitude characteristics of fusion alpha-particle driven ion cyclotron emission}
+\title[]{Doublet splitting of fusion alpha-particle driven ion cyclotron emission}
 
@@
 \bibitem{Dendy1995}
-Dendy R O, McClements K G, Lashmore-Davies C, Cottrell G A, Majeski R, and Cauffman S 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
+Dendy R O {\textit et al} 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
+%Dendy R O, McClements K G, Lashmore-Davies C, Cottrell G A, Majeski R, and Cauffman S 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
 %  , “Ion cyclotron emission due to collective instability of fusion prod-ucts and beam ions in TFTR and JET,”
@@
 \bibitem{Cottrell1993} 
-Cottrell G A, Bhatnagar V, Costa O D, Dendy R O, Jacquinot J, McClements K G, McCune D, Nave M, Smeulders P, and Start D 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
+Cottrell G A {\textit et al} 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
+%Cottrell G A, Bhatnagar V, Costa O D, Dendy R O, Jacquinot J, McClements K G, McCune D, Nave M, Smeulders P, and Start D 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
 % 
@@
 \bibitem{Ichimura2008}
-Ichimura M, Higaki H, Kakimoto S, Yamaguchi Y, Nemoto K, Katano M, Ishikawa M, Moriyama S, and Suzuki T 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
+Ichimura M {\textit et al} 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
+%Ichimura M, Higaki H, Kakimoto S, Yamaguchi Y, Nemoto K, Katano M, Ishikawa M, Moriyama S, and Suzuki T 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
 % , "Observation of spontaneously excited waves in the ion cyclotron frequency range on JT-60u,"
@@
 \bibitem{Chapman2017}
-Chapman B, Dendy R O, McClements K, Chapman S C, Yun G, Thatipamula S, and Kim M 2017 \textit{Nucl. Fusion} \textbf{57} 124004
+Chapman B {\textit et al} 2017 \textit{Nucl. Fusion} \textbf{57} 124004
+%Chapman B, Dendy R O, McClements K, Chapman S C, Yun G, Thatipamula S, and Kim M 2017 \textit{Nucl. Fusion} \textbf{57} 124004
 % , "Sub-microsecond temporal evolution of edge density during edge localized modes in KSTAR tokamak plasmas inferred from ion cyclotron emission,"
@@
 \bibitem{Chapman2018}
-Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027
+Chapman B {\textit et al} 2018 \textit{Nucl. Fusion} \textbf{58} 096027
+%Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027
 
@@
 \bibitem{McClements1996}
-McClements K G, Dendy R O, Lashmore-Davies C N, Cottrell G A, Cauffman S and Majeski R (1996) \textit{Phys. Plasmas} \textbf{3} 543
+McClements K G {\textit et al} 1996 \textit{Phys. Plasmas} \textbf{3} 543
+%McClements K G, Dendy R O, Lashmore-Davies C N, Cottrell G A, Cauffman S and Majeski R (1996) \textit{Phys. Plasmas} \textbf{3} 543
 
@@
 \bibitem{Reman2019}
-Reman B, Dendy R O, Akiyama T, Chapman S C, Cook J W S, Igami  H, Inagaki S, Saito K, and Yun G, 2019 \textit{Nucl. Fusion} \textbf{59} 096013
+Reman B {\textit et al} 2019 \textit{Nucl. Fusion} \textbf{59} 096013
+%Reman B, Dendy R O, Akiyama T, Chapman S C, Cook J W S, Igami  H, Inagaki S, Saito K, and Yun G, 2019 \textit{Nucl. Fusion} \textbf{59} 096013
 % “Interpreting observations of ion cyclotron emission from large helical device plasmas with beam-injected ion populations,”
@@
 \bibitem{Sharapov2014}
-Sharapov S E , Lilley M K, Akers R, Ayed N B, Cecconello M, Cook J W S, Cunningham G, and Verwichte E 2014 \textit{Phys. Plasmas} \textbf{21} 082501
+Sharapov S E {\textit et al} 2014 \textit{Phys. Plasmas} \textbf{21} 082501
+%Sharapov S E, Lilley M K, Akers R, Ayed N B, Cecconello M, Cook J W S, Cunningham G, and Verwichte E 2014 \textit{Phys. Plasmas} \textbf{21} 082501
 % , "Bi-directional Alfvén cyclotron instabilities in the mega-amp spherical tokamak"
@@
 \bibitem{Heidbrink2006}
-Heidbrink W W, Fredrickson E, Gorelenkov N N, Rhodes T, Zeeland  M A 2006 \textit{Nucl. Fusion} \textbf{46} 324
+Heidbrink W W {\textit et al} A 2006 \textit{Nucl. Fusion} \textbf{46} 324
+%Heidbrink W W, Fredrickson E, Gorelenkov N N, Rhodes T, Zeeland M A 2006 \textit{Nucl. Fusion} \textbf{46} 324
 % Observation of compressional Alfvén eigenmodes (CAE) in a conventional tokamak.
@@
 \bibitem{Fredrickson2019}
-Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
+Fredrickson E D {\textit et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
+%Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
 % "Emission in the ion cyclotron range of frequencies (ICE) on NSTX and NSTX-U",

### 2022-02-28 manuscript.tex
tweaked log scale vs linear text

@@
 The observations of ICE intensity are plotted on a log scale scaled to match
-the range of the linear MCI growth rates. This juxtaposition is reasonable since
-nonlinear PIC code simulations indicate that all growing modes stop when the
-fastest growing mode becomes nonlinear at time $\tau$. Under the assumption
-that all modes start growing in unison from identical seed amplitudes then the
-$log$ of the amplitude of a mode with growth rate $\gamma$ is proportional to
-$\gamma \tau$, allowing for a comparison with observations also on a log scale.
+the range of the linear MCI growth rates. Direct comparison of log-scale
+ICE observations and linear growth rates is justified to the extent that
+nonlinear PIC simulations indicate that the linearly unstable modes
+cease to grow once the fastest growing mode becomes
+nonlinear\cite{Cook2013}. Hence, the $log$ of a mode's amplitude is
+proportional to its growth rate under the assumption that
+that all modes start growing in unison from identical seed amplitudes.
 There are several points worth noting about this figure. First, the

### 2022-03-25 manuscript.tex
latest updates

@@
 
-The integrals over perpendicular velocity required for calculating the
+Integrals over perpendicular velocity required for calculating the
 dielectric tensor for Maxwellian distribution functions are given in terms of
 modified Bessel functions of the first kind\cite{Stix} for both electrons and
-ldeuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
-alpha-particles' ring distribution is calculated numerically via
+deuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
+alpha-particles' ring distribution are calculated numerically via
 Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK}
 library. An absolute and relative tolerance of $0$ and $10^{-8}$, respectively,
-on the value of the integrals is used.
+is used to confirm convergence of the integral values.
 
@@
 
-The summation over Bessel function indices in Eq. \ref{eq:dielectrictensor}
-continues until the relative change between the current and
+Bessel functions indices in Eq. \ref{eq:dielectrictensor} are summed outwards
+from 0 in both the positive and negative directions
+until the relative change between the current and
 previous iteration is less than a tolerance of $10^{-8}$. Root finding of Eq.
@@
 The observations of ICE intensity are plotted on a log scale scaled to match
-the range of the linear MCI growth rates. Direct comparison of log-scale
+the range of the linear MCI growth rates.
+Direct comparison of log-scale
 ICE observations and linear growth rates is justified to the extent that
-nonlinear PIC simulations indicate that the linearly unstable modes
+nonlinear PIC simulations indicate that the linearly unstable modes rising
+as $\exp(\gamma t)$, where $\gamma$ is a mode's growth rate,
 cease to grow once the fastest growing mode becomes
-nonlinear\cite{Cook2013}. Hence, the $log$ of a mode's amplitude is
-proportional to its growth rate under the assumption that
-that all modes start growing in unison from identical seed amplitudes.
+nonlinear\cite{Cook2013}.
+Therefore under the assumption that all modes start growing in unison from
+identical seed amplitudes then the log of a mode's amplitude is
+proportional to its growth rate.
 There are several points worth noting about this figure. First, the
@@
 $k_\parallel$ goes through zero combined with the change in sign of the Doppler
-shift of the real frequency as $k_\parallel$ changes sign. The effect is
+shift of the real frequency as $k_\parallel$ changes sign, which pushes the real
+frequency from one side of the harmonic to the other. The effect is
 larger for higher harmonics where the growth rate diminishes to zero for
@@
 
-This effect can be seen in Fig. \ref{fig:2D_imag_physicalpitch}, which is the
+Doublet splitting can be seen in Fig. \ref{fig:2D_imag_physicalpitch}, which is the
 counterpart figure to Fig. \ref{fig:2D_F12_physicalpitch} as Fig.

### 2022-05-08 manuscript.tex
updated for zenodo of code

@@
 rates are obtained from a dispersion relation solver,
-PlasmaDispersionRelations.jl, written in the Julia programming
+LinearMaxwellVlasov.jl\cite{LinearMaxwellVlasov.jl}, written in the Julia programming
 language\cite{Bezanson2017}. These results narrow the gap between analytical
@@
 tractable form. The computational capabilities of the software,
-PlasmaDispersionRelations.jl, used to generate the results further enable ICE as
+LinearMaxwellVlasov.jl, used to generate the results further enable ICE as
 a diagnostic for fusion plasmas since the code is capable of: calculating the
@@
 \bibitem{Dendy1995}
-Dendy R O {\textit et al} 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
+  Dendy R O \textit{et al} 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
 %Dendy R O, McClements K G, Lashmore-Davies C, Cottrell G A, Majeski R, and Cauffman S 1995 \textit{Nucl. Fusion} \textbf{35} 1733–1742
 %  , “Ion cyclotron emission due to collective instability of fusion prod-ucts and beam ions in TFTR and JET,”
+
+\bibitem{LinearMaxwellVlasov.jl}
+  Cook J W S, \textit{LinearMaxwellVlasov.jl (0.1.0)}, Zenodo. https://doi.org/10.5281/zenodo.6529288
 
@@
 \bibitem{Cottrell1993} 
-Cottrell G A {\textit et al} 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
+Cottrell G A \textit{et al} 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
 %Cottrell G A, Bhatnagar V, Costa O D, Dendy R O, Jacquinot J, McClements K G, McCune D, Nave M, Smeulders P, and Start D 1993 \textit{Nucl. Fusion} \textbf{33}, 31365–1387
@@
 \bibitem{Ichimura2008}
-Ichimura M {\textit et al} 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
+Ichimura M \textit{et al} 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
 %Ichimura M, Higaki H, Kakimoto S, Yamaguchi Y, Nemoto K, Katano M, Ishikawa M, Moriyama S, and Suzuki T 2008 \textit{Nucl. Fusion} \textbf{48}, 035012
@@
 \bibitem{Chapman2017}
-Chapman B {\textit et al} 2017 \textit{Nucl. Fusion} \textbf{57} 124004
+Chapman B \textit{et al} 2017 \textit{Nucl. Fusion} \textbf{57} 124004
 %Chapman B, Dendy R O, McClements K, Chapman S C, Yun G, Thatipamula S, and Kim M 2017 \textit{Nucl. Fusion} \textbf{57} 124004
@@
 \bibitem{Chapman2018}
-Chapman B {\textit et al} 2018 \textit{Nucl. Fusion} \textbf{58} 096027
+Chapman B \textit{et al} 2018 \textit{Nucl. Fusion} \textbf{58} 096027
 %Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027
@@
 \bibitem{McClements1996}
-McClements K G {\textit et al} 1996 \textit{Phys. Plasmas} \textbf{3} 543
+McClements K G \textit{et al} 1996 \textit{Phys. Plasmas} \textbf{3} 543
 %McClements K G, Dendy R O, Lashmore-Davies C N, Cottrell G A, Cauffman S and Majeski R (1996) \textit{Phys. Plasmas} \textbf{3} 543
@@
 \bibitem{Reman2019}
-Reman B {\textit et al} 2019 \textit{Nucl. Fusion} \textbf{59} 096013
+Reman B \textit{et al} 2019 \textit{Nucl. Fusion} \textbf{59} 096013
 %Reman B, Dendy R O, Akiyama T, Chapman S C, Cook J W S, Igami  H, Inagaki S, Saito K, and Yun G, 2019 \textit{Nucl. Fusion} \textbf{59} 096013
@@
 \bibitem{Sharapov2014}
-Sharapov S E {\textit et al} 2014 \textit{Phys. Plasmas} \textbf{21} 082501
+Sharapov S E \textit{et al} 2014 \textit{Phys. Plasmas} \textbf{21} 082501
 %Sharapov S E, Lilley M K, Akers R, Ayed N B, Cecconello M, Cook J W S, Cunningham G, and Verwichte E 2014 \textit{Phys. Plasmas} \textbf{21} 082501
@@
 \bibitem{Heidbrink2006}
-Heidbrink W W {\textit et al} A 2006 \textit{Nucl. Fusion} \textbf{46} 324
+Heidbrink W W \textit{et al} A 2006 \textit{Nucl. Fusion} \textbf{46} 324
 %Heidbrink W W, Fredrickson E, Gorelenkov N N, Rhodes T, Zeeland M A 2006 \textit{Nucl. Fusion} \textbf{46} 324
@@
 \bibitem{Fredrickson2019}
-Fredrickson E D {\textit et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
+Fredrickson E D \textit{et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
 %Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111

### 2022-05-08 ICE2D.jl

@@ -23,9 +23,9 @@ addprocs(nprocsadded)
 @everywhere using StaticArrays
 @everywhere using FastClosures
 @everywhere begin
-  using PlasmaDispersionRelations, LinearAlgebra, WindingNelderMead
+  using LinearMaxwellVlasov, LinearAlgebra, WindingNelderMead
 
-  mₑ = PlasmaDispersionRelations.mₑ
+  mₑ = LinearMaxwellVlasov.mₑ
   md = 2*1836*mₑ
   mα = 2*md
 
@@ -38,7 +38,7 @@ addprocs(nprocsadded)
   nd = n0 / (1.0 + 2*ξ)
   nα = ξ*nd
   @assert n0 ≈ 2*nα + nd
-  Va = sqrt(B0^2/PlasmaDispersionRelations.μ₀/nd/md)
+  Va = sqrt(B0^2/LinearMaxwellVlasov.μ₀/nd/md)
 
   Ωe = cyclotronfrequency(B0, mₑ, -1)
   Ωd = cyclotronfrequency(B0, md, 1)

### 2022-05-08 run.sh

@@ -3,9 +3,9 @@ source ~/.bashrc
 shopt -s expand_aliases
 
 # first arg is pitch, second arg is ring thermal spread as fraction of its speed
-julia1.7 -e 'using Pkg; Pkg.instantiate()'
-julia1.7 --proj ICE2D.jl -0.646 0.01 0.01 physical
-julia1.7 --proj ICE2D.jl  0.0   0.01 0.01 ring
+julia1.7 --proj=. -e 'using Pkg; Pkg.instantiate()'
+julia1.7 --proj=. ICE2D.jl -0.646 0.01 0.01 physical
+julia1.7 --proj=. ICE2D.jl  0.0   0.01 0.01 ring
 
 # now reduce size and improve images
 for j in physical ring

### 2022-07-17 manuscript.tex
first stab at amendments

@@
 alpha-particles and Maxwellian electrons and background thermal majority
-deuteron ions. Calculations are made for alpha-particles with ring-beams of two
-values of pitch-angle cosine: $u_\parallel=-0.646 u_0 \approx
--\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_{ICE}})) \simeq -1.1 V_A$, which is
-associated with the ion-cyclotron emission observed on JET Pulse No. 26148 for
-which $R_0=3.46$ m and
-$\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
-$R_{ICE}=4$ m is the major radius of the emission region, see
-Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
-$u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
-Fig. 1(a) of Ref \cite{Cook2013}. Here, $V_A$ is the Alfv{\'en} speed and $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is
-the speed of an alpha-particle with an energy of 3.6 MeV. The complex frequency
-solutions, $\omega$, along the fast Alfv{\'e}n branch are found on a grid of $(k_\parallel, k_\bot)$.
+deuteron ions.
 
-The plasma parameters approximate those found in the region indicated in Fig. 3
-of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
-signal of Fig. 2 in the same reference. The electrons and deuterons have
-temperatures of 1 keV and the electrons have a number density
-of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
-$\mathrm{T}$, which corresponds a major radius of $4$ m.
 The ring-beam distribution function, Eq. 1, is a separable
@@
 ring in $v_\perp$ given by Eq. 3.
-This distribution function is parameterised by the drift
+In the context of the fully kinetic Maxwell-Vlasov linear theory,
+which rests upon the assumption of
+an infinite homogeneous plasma and a straight uniform magnetic field,
+this distribution function is parameterised  by the drift
 velocity parallel to the magnetic field $u_\parallel$, the drift speed 
@@
 
+Although curvature and grad-B drift terms would be present in fast
+alpha-particle populations on JET, it is worth noting that they are both of
+order $\rho^*$ and have been neglected in these calculations where $u_\bot$ is
+always large.
+
+Calculations are made for alpha-particles with ring-beams of two
+values of pitch-angle cosine: $u_\parallel=-0.646 u_0 \approx
+-\cos(\arcsin(\sqrt{R_0 \sin(\psi_0)/R_{ICE}})) u_0 \simeq -1.1 V_A$, which is
+associated with the ion-cyclotron emission observed on JET Pulse No. 26148 for
+which $R_0=3.46$ m and
+$\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
+$R_{ICE}=4$ m is the major radius of the emission region, see
+Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
+$u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
+Fig. 1(a) of Ref \cite{Cook2013}. Here, $V_A$ is the Alfv{\'en} speed and $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is
+the speed of an alpha-particle with an energy of 3.6 MeV.
+
+The plasma parameters approximate those found in the region indicated in Fig. 3
+of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
+signal of Fig. 2 in the same reference. The electrons and deuterons have
+temperatures of 1 keV and the electrons have a number density
+of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
+$\mathrm{T}$, which corresponds a major radius of $4$ m.
+
 Solutions to the linearised Maxwell-Vlasov system of equations are given by the
 dispersion relation, Eq. \ref{eq:dettensor}, which is satisfied for sets of
-$(\omega,\vec{k})$ given the distribution functions of all the species present.
+$(\omega,\vec{k})$ given the distribution functions of all the species present
+along the fast Alfv{\'e}n branch  on a grid of $(k_\parallel, k_\bot)$.
 The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$, Eq. \ref{eq:dielectrictensor} is calculated for each species.
@@
 proportional to its growth rate.
+
+Although this simplication neglects possibly important effects, such as particle sink
+and source terms, and their influence on the nonlinear amplitudes of the
+observed ICE signals it is still possible to obtain important information
+regarding the signals' structure and overall trends.
+
 There are several points worth noting about this figure. First, the
@@
 of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
-signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
-Fig. 9 Dendy et al 1995) shows slices at constant propagation angle that when
+signal showing multiple doublet splittings; Fig. 3 of Ref. \cite{Dendy1994} (also
+Fig. 9 Ref. \cite{Dendy1995}) shows slices at constant propagation angle that when
 all superimposed point towards doublet splitting, albeit under-resolved in
@@
 No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
-the step change in signal strength from $\omega < 6$ to $\omega > 6$; the
+the step change in signal strength from $\omega < 6\Omega_i$ to $\omega > 6\Omega_i$; the
 doublet splitting of peaks at $\omega \approx 10\Omega_i$ in Fig.

### 2022-07-20 manuscript.tex
added refs to Liu and Ochoukov, multiplet splitting, drift resonances

@@
 this study focusses on edge-ICE, named for the region of plasma from which the signal
-arises, as opposed to core-ICE\cite{Ochoukov2018}.
+arises, as opposed to core-ICE\cite{Ochoukov2018}. Recent observations made with
+higher resolution equipment than those referred to on JET have
+shown multiplet splitting in NBI-driven spectra; from the core on
+ASDEX-Upgrade\cite{Ochoukov2019} and edge on EAST\cite{Lui2020}.
 
@@
 Other possible mechanisms for doublet splitting could be grad-B or curvature
-drifts\cite{Cottrell1993,Fulop1998}. Qualitatively, scattered points in Fig.
+drifts\cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
+of modifying the wave-particle resonance condition.
+Qualitatively, scattered points in Fig.
 \ref{fig:2D_F12_physicalpitch}(a) resembles the experimental observations given
@@
 
+\bibitem{Ochoukov2019}
+Ochoukov R \textit{et al} 2019 \textit{Nucl. Fusion} \textbf{59} 086032
+
+\bibitem{Lui2020}
+Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
+
 \bibitem{Fredrickson2019}

### 2022-07-23 manuscript.tex
further improvements based off referee comments

@@
 perturbed into a more stable state. Analysis shows that only higher cyclotron
-harmonics are destabilised\cite{McClements2015} to the MCI, however experimental
+harmonics are destabilised\cite{McClements2015} to the MCI under these
+conditions, however experimental
 observations clearly indicate power at theoretically stable lower
@@
 2D3V dynamics of ICE\cite{Carbajal2021}.
+
+Finer detailed structure has been obvserved in ICE spectra, such as doublet
+splitting of individual harmonics on JET\cite{Cottrell1993} and
+DIII-D\cite{Thome2018} and even
+multiplet splitting on TUMAN-3M\cite{Askinazi2018} on EAST\cite{Lui2020}.
+This paper highlights one underlying kinetic mechinism that produces
+doublet splitting due to Doppler shifted resonances. In general,
+splitting could be produced by grad-B or curvature
+drifts\cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
+of adding a Doppler shift to the wave-particle resonance condition.
 
@@
 
-Although curvature and grad-B drift terms would be present in fast
-alpha-particle populations on JET, it is worth noting that they are both of
-order $\rho^*$ and have been neglected in these calculations where $u_\bot$ is
-always large.
-
 Calculations are made for alpha-particles with ring-beams of two
@@
 $\mathrm{T}$, which corresponds a major radius of $4$ m.
+
+Although curvature and grad-B drift terms would be present in fast
+alpha-particle populations on JET, it is worth noting that they are both of
+order $\rho^*$ and have been neglected in these calculations where $u_\bot \gg
+\rho^* u_0$ is always the case.
 
@@
 k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed.
-Other possible mechanisms for doublet splitting could be grad-B or curvature
-drifts\cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
-of modifying the wave-particle resonance condition.
 Qualitatively, scattered points in Fig.
@@
     parallel wavenumber in units of $\Omega_i/V_A$ and the black trace shows
-    the re-scaled ICE signal from JET Pulse \#21648 digitised from Fig. 2 of
+    the re-scaled ICE signal from JET Pulse \#26148 digitised from Fig. 2 of
     Cottrell et al 1993 (with permission from IOP Publishing), (b) plots the solutions
@@
 
-It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou for useful discussions.
+It is a pleasure to thank R O Dendy and B Chapman-Oplopoiou for useful
+discussions.
 
@@
 
+\bibitem{Thome2018}
+Thome K E \textit{et al} 2018 \textit{Rev. Sci. Instrum.} \textbf{89} 10I102
+
+\bibitem{Askinazi2018}
+Askinazi L G  \textit{et al} 2018 \textit{Nucl. Fusion} \textbf{58} 082003
+
+\bibitem{Lui2020}
+Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
+
 \bibitem{Sharapov2014}
@@
 Ochoukov R \textit{et al} 2019 \textit{Nucl. Fusion} \textbf{59} 086032
-
-\bibitem{Lui2020}
-Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
 

### 2022-07-24 manuscript.tex
refer to Fig 5 and add high freq Va eqn

@@
 
-There are several points worth noting about this figure. First, the
+There are several points worth noting about Fig. \ref{fig:2D_F12_physicalpitch}.
+First, the
 growth rate is no longer linear with cyclotron harmonic number. Second,
@@
 attributed to the contributions from multiple propagation angles even though the
-lower harmonics are largely stable for forward travelling and perpendicular
-propagation angles.
+lower harmonics are largely stable for forward travelling and
+propagation angles that are close to perpendicular.
+
+Figure \ref{fig:2D_real_zeropitch}
+shows the smoothed real part of the solution for the case of zero pitch angle
+cosine and is some normalised by
+the high frequency expression for the fast Alfv{\'e}n wave set out in Eq.
+25 of Ref. \cite{Dendy1994} (also Eq. 11 of Ref. \cite{McClements1996}),
+
+\begin{equation}
+  \omega^2 = \frac{V_A^2}{2}\left[
+    k^2 + k_\parallel^2 + k^2k_\parallel^2 \frac{V_A^2}{\Omega_i^2}
+  + \sqrt{
+    \left(k^2 + k_\parallel^2 + k^2k_\parallel^2 \frac{V_A^2}{\Omega_i^2}\right)^2
+    - 4k^2 k_\parallel^2
+}
+  \right],
+\end{equation}
+
+\noindent where $k^2 = k_\parallel^2 + k_\perp^2$. The real part of the
+solutions for the case of non-zero pitch angle cosine is not discernably
+different.
 
@@
 No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
-the step change in signal strength from $\omega < 6\Omega_i$ to $\omega > 6\Omega_i$; the
+the step change in signal strength from $\omega < 6\Omega_i$
+to $\omega > 6\Omega_i$; the
 doublet splitting of peaks at $\omega \approx 10\Omega_i$ in Fig.

### 2022-07-26 manuscript.tex
post red-penning

@@
   in toroidal fusion plasmas is an important non-invasive, passive diagnostic
-  for current and next generation devices. The ability to model the ICE signals
+  for current and next generation devices. The ability to model ICE signals
   accurately is an essential step towards inferring the characteristics of
-  confined energetic alpha-particles or fast NBI ions and background plasma. In
-  this paper, the linear growth rates of the magnetoacoustic cyclotron
+  confined energetic alpha-particles or fast neutral beam ions and background plasma.
+  In this paper, the linear growth rates of the magnetoacoustic cyclotron
   instability, which is the leading explanation for ICE, are calculated in high
@@
   JET Pulse No. 26148. These calculations shed further light on the origin of
-  doublet-splitting of the cyclotron harmonics as well as the apparent need to
+  doublet-splitting of the cyclotron harmonics as well as the perceived need to
   invoke nonlinear interactions with 1D3V particle-in-cell studies to account
@@
 
-Observations of ion cyclotron emission (ICE) have been reported from magnetically confined plasmas first in TFTR\cite{Cauffman1995,Dendy1995} and JET\cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks PDX\cite{Heidbrink1994}, JT-60-U\cite{Ichimura2008}, DIII-D\cite{Heidbrink2011}, ASDEX\cite{DInca2014}, ASDEX-Upgrade\cite{Ochoukov2018}, EAST\cite{Liu2020}, and KSTAR\cite{Chapman2017,Chapman2018}.
-Stellarators\cite{Saito2013} and FRCs\cite{Nicks2020} have also exhibited this phenomenon.
-Energetic ion populations that give rise to ICE have their origins in fusion reactions in DT plasmas\cite{Cottrell2000, Cauffman1995}, ICRH\cite{Cottrell2000}, and neutral beam injection\cite{Ochoukov2018}.
-ICE arising from fusion $\alpha$-particles is a proposed route to further the understanding of burning plasmas in ITER\cite{Dendy2015,McClements2015}.
+Observations of ion cyclotron emission (ICE) have been reported from
+magnetically confined plasmas first in TFTR\cite{Cauffman1995,Dendy1995}
+and JET\cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks
+PDX\cite{Heidbrink1994}, JT-60-U\cite{Ichimura2008}, DIII-D\cite{Heidbrink2011},
+ASDEX\cite{DInca2014}, KSTAR\cite{Chapman2017,Chapman2018},
+ASDEX-Upgrade\cite{Ochoukov2018},
+TUMAN-3M\cite{Askinazi2018}, EAST\cite{Liu2020}. Stellarators\cite{Saito2013} and
+FRCs\cite{Nicks2020} have also exhibited this phenomenon. Energetic ion
+populations that give rise to ICE have their origins in fusion
+reactions in DT plasmas\cite{Cottrell2000, Cauffman1995},
+ICRH\cite{Cottrell2000}, and neutral beam injection\cite{Ochoukov2018}. ICE
+arising from fusion $\alpha$-particles is a proposed route to further
+the understanding of burning plasmas in ITER\cite{Dendy2015,McClements2015}.
 
@@
 $v_{th}$ are the beam velocity (the signed drift speed along the magnetic
-field), ring speed (the speed perpendicular to the magnetic field) and thermal
-velocity, respectively. Subsequent studies augmented this understanding via
-nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and
+field), ring speed (the particle speed perpendicular to the magnetic field)
+and thermal velocity, respectively. Subsequent studies augmented this
+understanding via nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and
 hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
@@
 inhomogeneous plasma and magnetic fields have shed light of the eigenmode
-structure of compressional Alfv{\'e}n eigenmodes, which like ICE are also fast
+structure of compressional Alfv{\'e}n eigenmodes, which as with ICE are also fast
 Alfv{\'e}n modes driven by energetic ions, in spherical
@@
 harmonic peaks in the fluctuations of the magnetic field was first explained
-somewhat by 1D PIC simulations. In these simulations the onset of nonlinearity
+somewhat by 1D PIC simulations\cite{Cook2013}.
+In these simulations the onset of nonlinearity
 appeared to imprint the amplitudes of each of the harmonics into the signal as
@@
 phase meaning that both the amount of free energy available for instability has
-decreased and that the unstable energy ion distribution function has been
+decreased and that the unstable ion distribution function has been
 perturbed into a more stable state. Analysis shows that only higher cyclotron
@@
 splitting of individual harmonics on JET\cite{Cottrell1993} and
-DIII-D\cite{Thome2018} and even
-multiplet splitting on TUMAN-3M\cite{Askinazi2018} on EAST\cite{Lui2020}.
+DIII-D\cite{Thome2018} and even multiplet splitting on TUMAN-3M\cite{Askinazi2018},
+ASDEX-Upgrade\cite{Ochoukov2019} and EAST\cite{Lui2020}.
 This paper highlights one underlying kinetic mechinism that produces
-doublet splitting due to Doppler shifted resonances. In general,
-splitting could be produced by grad-B or curvature
+doublet splitting due to Doppler shifted resonances and the
+intricate landscape of the MCI growth rate on the fast Alfv{\'e}n dispersion
+surface. In principle, splitting could be produced by grad-B or curvature
 drifts\cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
-of adding a Doppler shift to the wave-particle resonance condition.
+of adding a Doppler shift to the wave-particle resonance condition. Note that
+this study focusses on edge-ICE, named for the region of plasma from which the signal
+arises, as opposed to core-ICE\cite{Ochoukov2018, Liu2020}.
 
@@
 language\cite{Bezanson2017}. These results narrow the gap between analytical
-theory and expensive HPC simulations and can be run readily on desktop computers
-in a matter of minutes, further enabling research into this topic. Note that
-this study focusses on edge-ICE, named for the region of plasma from which the signal
-arises, as opposed to core-ICE\cite{Ochoukov2018}. Recent observations made with
-higher resolution equipment than those referred to on JET have
-shown multiplet splitting in NBI-driven spectra; from the core on
-ASDEX-Upgrade\cite{Ochoukov2019} and edge on EAST\cite{Lui2020}.
+theory and expensive HPC simulations and can be run on desktop computers
+in a matter of minutes, further enabling research into this topic.
 
-In section 2, the computational method is described for calculating the linear growth rates, which are presented for energetic ring-beam alpha-particles with non-zero perpendicular thermal spread in section 3.
+In section 2, the computational method is described for calculating
+the linear growth rates of the MCI, which are presented for energetic ring-beam
+alpha-particles with non-zero perpendicular thermal spread in section 3
+where a comparison is ade to the observations on JET..
 Results are summarised in section 4.
@@
 Solutions to the linearised Maxwell-Vlasov system of equations are found in
-parallel and perpendicular wavenumber space, $(k_\parallel, k_\bot)$, for a
+parallel and perpendicular wavenumber space, $(k_\bot, k_\parallel)$, for a
 ring-beam distribution of energetic minority
 alpha-particles and Maxwellian electrons and background thermal majority
-deuteron ions.
-
-The ring-beam distribution function, Eq. 1, is a separable
+deuteron ions. The ring-beam distribution function, Eq. 1, is a separable
 function of the beam in $v_\parallel$ given by Eq. 2 and the
@@
 perpendicular to the magnetic
-field $u_{\perp}$, and the thermal velocity in both directions $v_{th}$. The
+field $u_{\perp}$, and the isotropric thermal velocity $v_{th}$. The
 thermal spread of the alpha particles' distribution function 
 $v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$. The number density
-ratio of alpha-particles to deuterons is $1.5\times10^{-4}$.
+ratio of alpha-particles to the majority fuel deuterons is $1.5\times10^{-4}$.
 
@@
   th} u_\bot (1 - \mathrm{erf}(-\frac{u_\bot}{v_{\perp,th}})) + \exp(-u_\bot^2 /
-  v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}\\
+  v_{\perp, th}^2) \frac{v_{\perp,th}^2}{2}\right)}
 \end{align}
@@
 $\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
-$R_{ICE}=4$ m is the major radius of the emission region, see
+$R_{ICE}=4$ m is the major radius of the emission region; see
 Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
-$u_\parallel=0 u_0$ representing a ring distribution with no parallel drift, see
+$(u_\parallel=0 u_0, u_\bot=u_0)$ representing a ring distribution with
+no parallel drift, see
 Fig. 1(a) of Ref \cite{Cook2013}. Here, $V_A$ is the Alfv{\'en} speed and $u_0=\sqrt{u_\parallel^2 + u_\perp^2}$ is
@@
 
-The plasma parameters approximate those found in the region indicated in Fig. 3
-of Ref. \cite{Cottrell1993}, which are associated with the continuous trace
-signal of Fig. 2 in the same reference. The electrons and deuterons have
+The plasma parameters approximate those found at the outer mid-plane edge
+$(R=4\, \mathrm{m}, Z=0\, \mathrm{m})$ where the electrons and deuterons have
 temperatures of 1 keV and the electrons have a number density
 of $1.7\times 10^{19}$ $\mathrm{m^{-3}}$. The magnetic field is $2.07$
-$\mathrm{T}$, which corresponds a major radius of $4$ m.
+$\mathrm{T}$ at this major radius.
 
 Although curvature and grad-B drift terms would be present in fast
-alpha-particle populations on JET, it is worth noting that they are both of
+alpha-particle populations on JET and these drifts would alter the wave-particle
+resonance condition, it is worth noting that they are both of
 order $\rho^*$ and have been neglected in these calculations where $u_\bot \gg
@@
 along the fast Alfv{\'e}n branch  on a grid of $(k_\parallel, k_\bot)$.
-The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$, Eq. \ref{eq:dielectrictensor} is calculated for each species.
+The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$ given by Eq.
+\ref{eq:dielectrictensor} is calculated for each species, $s$,
 
@@
 deuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
-alpha-particles' ring distribution are calculated numerically via
+alpha-particles' ring distribution given by Eq. 3 are calculated numerically via
 Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK}
@@
 is used to confirm convergence of the integral values.
-
 Since all the parallel distribution functions for all species are Maxwellians,
@@
 
-Bessel functions indices in Eq. \ref{eq:dielectrictensor} are summed outwards
+Bessel function indices, $n$, in Eq. \ref{eq:dielectrictensor} are summed outwards
 from 0 in both the positive and negative directions
 until the relative change between the current and
-previous iteration is less than a tolerance of $10^{-8}$. Root finding of Eq.
-\ref{eq:dettensor} in complex frequency plane is performed by an amalgamation of the
-Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method,
-whereby if the values at the nodes of the Nelder-Mead simplex, which in this
+previous iteration is less than a tolerance of $10^{-8}$. 
+Root finding of Eq. \ref{eq:dettensor} in the complex frequency plane
+for a given $(k_\bot, k_\parallel)$ is performed by an amalgamation of the
+Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method.
+If the complex values at the nodes of the Nelder-Mead simplex, which in this
 case is just a triangle, have a winding number
@@
 The child simplex that contains the root is used for the next
-iteration. The optimisation iteration procedure finishes when the real extent of
-the simplex is less than $10^{-4}\, \Omega_i$ and the imaginary extent is less
-than $10^{-5}\,\Omega_i$, where $\Omega_i$ is the cyclotron frequency of both
+iteration. The optimisation iteration procedure finishes successfully when
+a simplex that encircles a root is small enough such that its extent in real
+(imaginary) frequency is less than $10^{-4}\, \Omega_i$ $(10^{-5}\, \Omega_i)$
+, where $\Omega_i$ is the cyclotron frequency of both
 the alpha-particles and deuterons since they share a charge-mass ratio.
 This approach is found to be adequately robust and
-faster than the Nelder-Mead method or winding number approach used in isolation
-of one another. Each 512 by 512 scan in $(k_\perp, k_\parallel)$ takes between 5
-and 6 minutes on a laptop.
+faster than either the Nelder-Mead method or winding number approach used
+separately. Each 512 by 512 scan in $(k_\perp, k_\parallel)$ takes between 5
+and 6 minutes on a current laptop.
 
@@
 in the $(\omega,k_\parallel)$ plane where shading instead indicates growth
-rates. The simplicity of the consecutive unstable cyclotron harmonics visible in
+rates in units of $\Omega_i$.
+The simplicity of the consecutive unstable cyclotron harmonics visible in
 panel (a) betrays the complexity of the structure in the $(\omega,k_\parallel)$
@@
     as in panel (a) in the $(\omega,k_\parallel)$ plane where shading indicates 
-    growth rate. The complex structure of the growth rates that exists in
+    growth rate in units of $\Omega_i$.
+    The complex structure of the growth rates that exists in
     wavenumber space is projected into sharp harmonics on the frequency axis.}
@@
 strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
-of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
-signal showing multiple doublet splittings; Fig. 3 of Ref. \cite{Dendy1994} (also
-Fig. 9 Ref. \cite{Dendy1995}) shows slices at constant propagation angle that when
-all superimposed point towards doublet splitting, albeit under-resolved in
-wavenumber space. The following describes how these three features are
+of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Ref.
+\cite{Cottrell1993} for a single
+signal showing multiple doublet splittings.
+The following describes how these three features are
 connected. Doublet splitting arises due to the dip in growth rate along the 
@@
 perpendicular propagation.
+This is hinted at by  Fig. 3 of Ref. \cite{Dendy1994} (also
+Fig. 9 Ref. \cite{Dendy1995}), which shows growth rates at slices of
+propagation angle that when all superimposed point towards doublet splitting,
+albeit under-resolved in wavenumber space.
 
@@
 larger and the harmonics are less easily identifiable. These features cannot be
-be recreated with a 1D calculation for these distributions functions. Note that
-Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes more
-clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is
+be recreated with a 1D calculation for these distributions functions
+since they arise from contributions from a range of propagation angles.
+Note that Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes
+more clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is
 characterised by narrow regions of instability near Doppler shifted cyclotron
@@
 negative parallel wavenumbers. The higher frequency behaviour in the region
-$\omega\geq 7\Omega_i$ is dominated by broader regions of instability in
-frequency and wavenumber near to Doppler shifted cyclotron harmonics, which have
-a complex structure in wavenumber space. Modes numbering 1, 2, 6, 7, 8, 12, and
+$k_\bot \geq 7\Omega_i/V_A$ is dominated by broader bands of instability
+near to Doppler shifted cyclotron harmonics, which have
+an intricate structure in wavenumber space. Modes numbering 1, 2, 6, 7, 8, 12, and
 13 are stable when propagating exactly perpendicular whilst obliquely
@@
 associated with the $15^{th}$ cyclotron harmonic near $k_\parallel \simeq
--\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 14 \Omega_i/V_A$, i.e. a propagation
-angle of $93^\circ$. Furthermore, the breadth of the harmonics can be
+-\frac{3}{4} \Omega_i/V_A, k_\perp \simeq 13 \Omega_i/V_A$, i.e. a propagation
+angle of $\sim 93^\circ$. Furthermore, the breadth of the harmonics can be
 attributed to the contributions from multiple propagation angles even though the
-lower harmonics are largely stable for forward travelling and
+lower harmonics are largely stable for forward travelling modes and
 propagation angles that are close to perpendicular.
@@
 Figure \ref{fig:2D_real_zeropitch}
-shows the smoothed real part of the solution for the case of zero pitch angle
-cosine and is some normalised by
+shows the smoothed real frequency part of the solution for the case of
+zero pitch angle cosine and is some normalised by
 the high frequency expression for the fast Alfv{\'e}n wave set out in Eq.
@@
   \right],
+\label{eqn:fastalfven}
 \end{equation}
 
-\noindent where $k^2 = k_\parallel^2 + k_\perp^2$. The real part of the
-solutions for the case of non-zero pitch angle cosine is not discernably
-different.
+\noindent where $k^2 = k_\parallel^2 + k_\perp^2$. There is no discernable difference for the non-zero pitch angle cosine case.
 
@@
     case with parallel drift $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$.
-    Panel (a) plots growth rate against frequency where shading indicates
+    Upper panel (a) plots growth rate against frequency where shading indicates
     parallel wavenumber in units of $\Omega_i/V_A$ and the black trace shows
     the re-scaled ICE signal from JET Pulse \#26148 digitised from Fig. 2 of
-    Cottrell et al 1993 (with permission from IOP Publishing), (b) plots the solutions
+    Cottrell et al 1993 (with permission from IOP Publishing). Lower panel (b)
+    plots the solutions
     in the $(\omega,k_\parallel)$ plane where shading indicates growth rate. Note
@@
     the shading indicates the real part of the smoothed solution normalised by
-    the high frequency expression for the fast Alfv{\'e}n wave set out in Eq. 25
-    of Ref. \cite{Dendy1994}, which is also Eq. 11 in Ref.
-    \cite{McClements1996}. The differences in real frequency for the two sets of
+    the high frequency expression for the fast Alfv{\'e}n wave set out in
+    Eq. \label{eqn:fastalfven} (Eq. 25 of Ref. \cite{Dendy1994},
+    which is also Eq. 11 in Ref. \cite{McClements1996}).
+    The differences in real frequency for the two sets of
     calculations, a pitch-angle cosine of $-0.646$ and zero, are
@@
 No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
-the step change in signal strength from $\omega < 6\Omega_i$
+the step change in signal strength from $\omega \leq 6\Omega_i$
 to $\omega > 6\Omega_i$; the
@@
 for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron
-frequency, which makes it difficult to determine which parts of the signal come
-from which cyclotron harmonic numbers. Further effort is required to account for
+frequency making it difficult to determine which parts of the signal come
+from which cyclotron harmonic numbers, which has important implications for
+determining the major radius of origin of ICE.
+Further effort is required to account for
 the extra doublets visible in the continuous trace associated with JET Pulse No.
@@
 physics\cite{Chapman2018} may also play a part.
+
 There are several important features missing from these calculations. Among them
 are the effects of the spatial inhomogeniety of the magnetic field and background
-plasma, and the quantisation associatied with toroidal periodicity. These
+plasma giving rise to drifts,
+and the quantisation associatied with toroidal periodicity. These
 effects may create an eigenmode structure, not accounted for herein,
@@
 
-This work showcases a new high resolution tool for studying ICE and captures the
-fine detail of ICE signals, which is challenging to capture with nonlinear PIC
+This work showcases a new high resolution tool for studying ICE and captures
+some of the fine detail of ICE signals, which is challenging to capture
+with nonlinear PIC
 code simulations since they suffer from low resolution in frequency and
@@
 
-\bibitem{LinearMaxwellVlasov.jl}
-  Cook J W S, \textit{LinearMaxwellVlasov.jl (0.1.0)}, Zenodo. https://doi.org/10.5281/zenodo.6529288
-
 \bibitem{Cottrell1988}
@@
 
-\bibitem{Ochoukov2018}
-Ochoukov R et al 2018 \textit{Rev. Sci. Instrum.} 2018 \textbf{89}, 10J101
-% , "Observations of core ion cyclotron emission on ASDEX Upgrade tokamak"
-% , Bobkov V, Chapman B, Dendy RO and Dunne,M  and Faugel, H  and García-Muñoz,M  and Geiger,B  and Hennequin,P  and McClements,K G  and Moseev,D  and Nielsen,S  and Rasmussen,J  and Schneider,P  and Weiland,M  and Noterdaeme,J-M
-
-\bibitem{Liu2020}
-Liu L et al 2020  \textit{Nucl. Fusion} \textbf{60} 044002
-% "Ion cyclotron emission driven by deuterium neu- tral beam injection and core fusion reaction ions in EAST,"
-% Zhang X, Zhu Y, Qin C, Y Zhao, S Yuan, Y Mao, M Li, K Zhang, J Cheng, L Ai, and Y Cheng
-
 \bibitem{Chapman2017}
@@
 %Chapman B, Dendy R O, Chapman S C, McClements K G, Yun G S and Thatipamula S G 2018 \textit{Nucl. Fusion} \textbf{58} 096027
+
+\bibitem{Ochoukov2018}
+Ochoukov R et al 2018 \textit{Rev. Sci. Instrum.} 2018 \textbf{89}, 10J101
+% , "Observations of core ion cyclotron emission on ASDEX Upgrade tokamak"
+% , Bobkov V, Chapman B, Dendy RO and Dunne,M  and Faugel, H  and García-Muñoz,M  and Geiger,B  and Hennequin,P  and McClements,K G  and Moseev,D  and Nielsen,S  and Rasmussen,J  and Schneider,P  and Weiland,M  and Noterdaeme,J-M
+
+\bibitem{Askinazi2018}
+Askinazi L G  \textit{et al} 2018 \textit{Nucl. Fusion} \textbf{58} 082003
+
+\bibitem{Liu2020}
+Liu L et al 2020  \textit{Nucl. Fusion} \textbf{60} 044002
+% "Ion cyclotron emission driven by deuterium neu- tral beam injection and core fusion reaction ions in EAST,"
+% Zhang X, Zhu Y, Qin C, Y Zhao, S Yuan, Y Mao, M Li, K Zhang, J Cheng, L Ai, and Y Cheng
 
@@
 
-\bibitem{Thome2018}
-Thome K E \textit{et al} 2018 \textit{Rev. Sci. Instrum.} \textbf{89} 10I102
-
-\bibitem{Askinazi2018}
-Askinazi L G  \textit{et al} 2018 \textit{Nucl. Fusion} \textbf{58} 082003
-
-\bibitem{Lui2020}
-Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
-
 \bibitem{Sharapov2014}
@@
 
+\bibitem{Thome2018}
+Thome K E \textit{et al} 2018 \textit{Rev. Sci. Instrum.} \textbf{89} 10I102
+
+\bibitem{Ochoukov2019}
+Ochoukov R \textit{et al} 2019 \textit{Nucl. Fusion} \textbf{59} 086032
+
+\bibitem{Lui2020}
+Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
+
+\bibitem{LinearMaxwellVlasov.jl}
+  Cook J W S, \textit{LinearMaxwellVlasov.jl (0.1.0)}, Zenodo. https://doi.org/10.5281/zenodo.6529288
+
 \bibitem{Bezanson2017}
@@
 % Julia: A fresh approach to numerical computing,
-
-\bibitem{Ochoukov2019}
-Ochoukov R \textit{et al} 2019 \textit{Nucl. Fusion} \textbf{59} 086032
 

### 2022-07-28 manuscript.tex
careful re-reading improvements

@@
 Observations of ion cyclotron emission (ICE) have been reported from
-magnetically confined plasmas first in TFTR\cite{Cauffman1995,Dendy1995}
-and JET\cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks
-PDX\cite{Heidbrink1994}, JT-60-U\cite{Ichimura2008}, DIII-D\cite{Heidbrink2011},
-ASDEX\cite{DInca2014}, KSTAR\cite{Chapman2017,Chapman2018},
-ASDEX-Upgrade\cite{Ochoukov2018},
-TUMAN-3M\cite{Askinazi2018}, EAST\cite{Liu2020}. Stellarators\cite{Saito2013} and
-FRCs\cite{Nicks2020} have also exhibited this phenomenon. Energetic ion
+magnetically confined plasmas first in TFTR \cite{Cauffman1995,Dendy1995}
+and JET \cite{Cottrell1988,Cottrell1993,Dendy1995}, followed by other tokamaks
+PDX \cite{Heidbrink1994}, JT-60-U \cite{Ichimura2008}, DIII-D \cite{Heidbrink2011},
+ASDEX \cite{DInca2014}, KSTAR \cite{Chapman2017,Chapman2018},
+ASDEX-Upgrade \cite{Ochoukov2018},
+TUMAN-3M \cite{Askinazi2018}, both NSTX and NSTX-U \cite{Fredrickson2019},
+and EAST \cite{Liu2020}. Stellarators \cite{Saito2013} and
+FRCs \cite{Nicks2020} have also exhibited this phenomenon. Energetic ion
 populations that give rise to ICE have their origins in fusion
-reactions in DT plasmas\cite{Cottrell2000, Cauffman1995},
-ICRH\cite{Cottrell2000}, and neutral beam injection\cite{Ochoukov2018}. ICE
+reactions in DT plasmas \cite{Cottrell2000, Cauffman1995},
+ICRH \cite{Cottrell2000}, and neutral beam injection \cite{Ochoukov2018}. ICE
 arising from fusion $\alpha$-particles is a proposed route to further
-the understanding of burning plasmas in ITER\cite{Dendy2015,McClements2015}.
+the understanding of burning plasmas in ITER \cite{Dendy2015,McClements2015}.
 
 The foremost explanation of the mechanism responsible for ICE is the
-magnetoacoustic cyclotron instability (MCI)\cite{Belikov1976}, which arises from
+magnetoacoustic cyclotron instability (MCI) \cite{Belikov1976}, which arises from
 the free energy present in energetic ions with inverted distributions in
-velocity space. Theoretical understanding of ICE is based in analytical
-solutions to the linearised Maxwell-Vlasov system of equations\cite{Stix} where
-the energetic ion species was modelled as ring-beam for which $f(v_\parallel,
+velocity space. Theoretical understanding of ICE is based on analytical
+solutions to the linearised Maxwell-Vlasov system of equations \cite{Stix} where
+the energetic ion species is modelled as ring-beam for which $f(v_\parallel,
 v_\perp) \propto \exp(-(v_\parallel-u_\parallel)^2/ v_{th}^2)\delta(v_\perp -
@@
 $v_{th}$ are the beam velocity (the signed drift speed along the magnetic
-field), ring speed (the particle speed perpendicular to the magnetic field)
+field), ring speed (the speed of the shifted centre of the inverted population
+perpendicular to the magnetic field)
 and thermal velocity, respectively. Subsequent studies augmented this
-understanding via nonlinear fully-kinetic\cite{Cook2013,Chapman2017} and
-hybrid\cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
-spatial dimension and three vector components for velocities and electromagnetic
-field (i.e. 1D3V), and recently 2D3V hybrid-PIC simulations\cite{Carbajal2021}. These
+understanding via nonlinear fully-kinetic \cite{Cook2013,Chapman2017} and
+hybrid \cite{Carbajal2014,Reman2019} particle-in-cell (PIC) simulations in one
+spatial dimension with three vector components for velocities and electromagnetic
+field (i.e. 1D3V), and recently 2D3V hybrid-PIC simulations \cite{Carbajal2021}. These
 computational approaches operate under similar simplifications as the analytical
@@
 Alfv{\'e}n modes driven by energetic ions, in spherical
-tokamaks\cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio
+tokamaks \cite{Sharapov2014, Gorelenkov2016} and conventional aspect ratio
 tokamaks \cite{Kolesnichenko2000, Fulop2000, Heidbrink2006}.
 
-The effectiveness of analytical theory to reproduce the observed cyclotron
+The effectiveness of linear theory to reproduce the observed cyclotron
 harmonic peaks in the fluctuations of the magnetic field was first explained
-somewhat by 1D PIC simulations\cite{Cook2013}.
+somewhat by 1D PIC simulations \cite{Cook2013}.
 In these simulations the onset of nonlinearity
@@
 perturbed into a more stable state. Analysis shows that only higher cyclotron
-harmonics are destabilised\cite{McClements2015} to the MCI under these
-conditions, however experimental
+harmonics are destabilised \cite{McClements2015} to the MCI under these
+particular conditions, however experimental
 observations clearly indicate power at theoretically stable lower
-harmonics\cite{Cottrell1988}. In a further step towards accurate recreation of
-experimental observations, nonlinear phases of hybrid\cite{Carbajal2014} and
-fully kinetic\cite{Cook2013,Chapman2017} PIC simulations went on to show that
+harmonics \cite{Cottrell1988}. In a further step towards accurate recreation of
+experimental observations, nonlinear phases of hybrid \cite{Carbajal2014} and
+fully kinetic \cite{Cook2013,Chapman2017} PIC simulations went on to show that
 wave-wave coupling was responsible for raising the amplitude of lower cyclotron
 harmonics. Most recently, hybrid 2D simulations have begun to shed light on the
-2D3V dynamics of ICE\cite{Carbajal2021}.
+2D3V dynamics of ICE \cite{Carbajal2021}.
 
 Finer detailed structure has been obvserved in ICE spectra, such as doublet
-splitting of individual harmonics on JET\cite{Cottrell1993} and
-DIII-D\cite{Thome2018} and even multiplet splitting on TUMAN-3M\cite{Askinazi2018},
-ASDEX-Upgrade\cite{Ochoukov2019} and EAST\cite{Lui2020}.
+splitting of individual harmonics on JET \cite{Cottrell1993} and
+DIII-D \cite{Thome2018} and even multiplet splitting on TUMAN-3M \cite{Askinazi2018},
+ASDEX-Upgrade \cite{Ochoukov2019} and EAST \cite{Liu2020}, which has been made
+possible by advances in resolution of diagnostic techniques.
 This paper highlights one underlying kinetic mechinism that produces
@@
 intricate landscape of the MCI growth rate on the fast Alfv{\'e}n dispersion
-surface. In principle, splitting could be produced by grad-B or curvature
-drifts\cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
+surface in $(k_\bot, k_\parallel)$ space.
+In principle, splitting could be produced by grad-B or curvature
+drifts \cite{Cottrell1993,Fulop1998} or potentially any mechanism that is capable
 of adding a Doppler shift to the wave-particle resonance condition. Note that
 this study focusses on edge-ICE, named for the region of plasma from which the signal
-arises, as opposed to core-ICE\cite{Ochoukov2018, Liu2020}.
+arises, as opposed to core-ICE \cite{Ochoukov2018, Liu2020}.
 
 To better enable ICE as a diagnostic for confined fast ions, it is necessary to
-further develop our understanding with a range of tools. In this work, I present
+further develop our understanding with a range of tools. This work presents
 the growth rates in 2D wavenumber space of the MCI under the conditions of
-outer mid-plane of the JET PTE\cite{Cottrell1993} Pulse No. 26148. These growth
+outer mid-plane of the JET PTE \cite{Cottrell1993} Pulse No. 26148. These growth
 rates are obtained from a dispersion relation solver,
-LinearMaxwellVlasov.jl\cite{LinearMaxwellVlasov.jl}, written in the Julia programming
-language\cite{Bezanson2017}. These results narrow the gap between analytical
+LinearMaxwellVlasov.jl \cite{LinearMaxwellVlasov.jl}, written in the Julia programming
+language \cite{Bezanson2017}. These results narrow the gap between analytical
 theory and expensive HPC simulations and can be run on desktop computers
@@
 alpha-particles with non-zero perpendicular thermal spread in section 3
-where a comparison is ade to the observations on JET..
+where a comparison is made to the observations on JET.
 Results are summarised in section 4.
@@
 Solutions to the linearised Maxwell-Vlasov system of equations are found in
-parallel and perpendicular wavenumber space, $(k_\bot, k_\parallel)$, for a
+perpendicular and parallel wavenumber space, $(k_\bot, k_\parallel)$, for a
 ring-beam distribution of energetic minority
-alpha-particles and Maxwellian electrons and background thermal majority
+alpha-particles, Maxwellian electrons and background thermal majority
 deuteron ions. The ring-beam distribution function, Eq. 1, is a separable
@@
 this distribution function is parameterised by the drift
-velocity parallel to the magnetic field $u_\parallel$, the drift speed 
-perpendicular to the magnetic
+velocity parallel to the magnetic field $u_\parallel$, the speed of the
+Maxwellian distribution perpendicular to the magnetic
 field $u_{\perp}$, and the isotropric thermal velocity $v_{th}$. The
-thermal spread of the alpha particles' distribution function 
+thermal spread of the alpha-particles' distribution function 
 $v_{th}=v_{\parallel,th}=v_{\perp,th}=u_0/100$. The number density
@@
 resonance condition, it is worth noting that they are both of
-order $\rho^*$ and have been neglected in these calculations where $u_\bot \gg
+order of the ratio of the Larmor radius to the machine size, $\rho^*$,
+and have been neglected in these calculations where $u_\bot \gg
 \rho^* u_0$ is always the case.
@@
 dispersion relation, Eq. \ref{eq:dettensor}, which is satisfied for sets of
-$(\omega,\vec{k})$ given the distribution functions of all the species present
+$(\omega,\vec{k})$ 
 along the fast Alfv{\'e}n branch  on a grid of $(k_\parallel, k_\bot)$.
-The hot dielectric tensor\cite{Stix}, $\vec{\vec{\epsilon}}$ given by Eq.
-\ref{eq:dielectrictensor} is calculated for each species, $s$,
+The hot dielectric tensor \cite{Stix} $\vec{\vec{\epsilon}}$, given by Eq.
+\ref{eq:dielectrictensor}, is calculated for each species, $s$,
 
@@
 dielectric tensor for Maxwellian distribution functions are given in terms of
-modified Bessel functions of the first kind\cite{Stix} for both electrons and
+modified Bessel functions of the first kind \cite{Stix} for both electrons and
 deuterons i.e. when $u_\perp =0$. The same perpendicular integrals over the
 alpha-particles' ring distribution given by Eq. 3 are calculated numerically via
-Gauss-Kronrod\cite{Kronrod1965} quadrature using the GaussGK.jl\cite{QuadGK}
+Gauss-Kronrod \cite{Kronrod1965} quadrature using the GaussGK.jl \cite{QuadGK}
 library. An absolute and relative tolerance of $0$ and $10^{-8}$, respectively,
@@
 
-Bessel function indices, $n$, in Eq. \ref{eq:dielectrictensor} are summed outwards
-from 0 in both the positive and negative directions
-until the relative change between the current and
+Bessel function indices, $n$, in Eq. \ref{eq:dielectrictensor} are incremented
+outwards from 0 in both the positive and negative directions
+until the relative change between the current summand and that of the
 previous iteration is less than a tolerance of $10^{-8}$. 
 Root finding of Eq. \ref{eq:dettensor} in the complex frequency plane
-for a given $(k_\bot, k_\parallel)$ is performed by an amalgamation of the
-Nelder-Mead\cite{Nelder1965} optimisation method with the winding number method.
+for a given $(k_\bot, k_\parallel)$-pair is performed by an amalgamation of the
+Nelder-Mead \cite{Nelder1965} optimisation method with the winding number method.
 If the complex values at the nodes of the Nelder-Mead simplex, which in this
@@
 iteration. The optimisation iteration procedure finishes successfully when
-a simplex that encircles a root is small enough such that its extent in real
-(imaginary) frequency is less than $10^{-4}\, \Omega_i$ $(10^{-5}\, \Omega_i)$
-, where $\Omega_i$ is the cyclotron frequency of both
+the simplex that encircles a root is small enough such that its extent in real
+(imaginary) frequency is less than $10^{-4}\, \Omega_i$ $(10^{-5}\, \Omega_i)$,
+where $\Omega_i$ is the cyclotron frequency of both
 the alpha-particles and deuterons since they share a charge-mass ratio.
@@
 in the $(\omega,k_\parallel)$ plane where shading instead indicates growth
-rates in units of $\Omega_i$.
+rate in units of $\Omega_i$.
 The simplicity of the consecutive unstable cyclotron harmonics visible in
@@
 axis, which loses the structure and produces narrow harmonics. Figure
-\ref{fig:2D_imag_zeropitch} gives $(k_\perp, k_\parallel)$ space view of the
+\ref{fig:2D_imag_zeropitch} gives a $(k_\perp, k_\parallel)$ space view of the
 same data as Fig. \ref{fig:2D_F12_zeropitch} as well as damped and
-normal modes, except shading indicates the growth rates. Overlaid
+normal modes, except shading indicates growth rate. Overlaid
 contours indicate the real frequency in units of ion cyclotron harmonics. Only
 narrow regions around the cyclotron harmonics are unstable. Overlaid straight
-traces indicate propagation angle. No Doppler shift is present due to the lack
+traces show propagation angle. No Doppler shift is present due to the lack
 of parallel drift in the alpha-particle ring-beam. The growth rate increases
 approximately linearly with cyclotron harmonic number, which is consistent with
-the literature\cite{Dendy1994}. Together these figures serve to introduce a 2D
+the literature \cite{Dendy1994}. Together these figures serve to introduce a 2D
 view (growth rates in $(k_\perp, k_\parallel)$ space) of traditionally 1D data
@@
 Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physically more
-realistic pitch-angle
-case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
+realistic pitch case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
 \ref{fig:2D_F12_zeropitch} except there is an additional black trace in the
@@
 trace from Fig. 2 in Ref. \cite{Cottrell1993}.
-The observations of ICE intensity are plotted on a log scale scaled to match
+These observations of ICE intensity are plotted on a log scale scaled to match
 the range of the linear MCI growth rates.
@@
 cease to grow once the fastest growing mode becomes
-nonlinear\cite{Cook2013}.
+nonlinear \cite{Cook2013}.
 Therefore under the assumption that all modes start growing in unison from
@@
 since they arise from contributions from a range of propagation angles.
-Note that Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes
+Fig. \ref{fig:2D_imag_physicalpitch} shows these two different regimes
 more clearly. The lower frequency behaviour, where $k_\perp < 7 \Omega_i/V_A$, is
@@
 near to Doppler shifted cyclotron harmonics, which have
-an intricate structure in wavenumber space. Modes numbering 1, 2, 6, 7, 8, 12, and
-13 are stable when propagating exactly perpendicular whilst obliquely
+an intricate structure in wavenumber space. Modes numbering 1-5 and
+14 are stable when propagating exactly perpendicular whilst obliquely
 propagating modes remain unstable. The magnitude of the Doppler shift of
@@
 Figure \ref{fig:2D_real_zeropitch}
-shows the smoothed real frequency part of the solution for the case of
-zero pitch angle cosine and is some normalised by
+shows the smoothed real frequency part of the solution, as depicted
+by contours on Figs. \ref{fig:2D_imag_zeropitch} and
+\ref{fig:2D_imag_physicalpitch}, for the case of
+zero pitch angle cosine and is normalised by
 the high frequency expression for the fast Alfv{\'e}n wave set out in Eq.
@@
 
-\noindent where $k^2 = k_\parallel^2 + k_\perp^2$. There is no discernable difference for the non-zero pitch angle cosine case.
+\noindent where $k^2 = k_\parallel^2 + k_\perp^2$. There is no discernable
+difference for the non-zero pitch case.
 
@@
     \caption{(colour online) As in Fig. \ref{fig:2D_imag_zeropitch}, except for
-    a ring-beam of alpha particles with a more realistic pitch-angle value
+    a ring-beam of alpha-particles with a more realistic pitch-angle value
     $u_{\parallel} = -0.646 u_0 \simeq -1.1 V_A$. The non-zero parallel drift
@@
 $(k_\perp, k_\parallel)$ calculations revisit ICE observations from JET Pulse
-No. 26148\cite{Cottrell1993,Dendy1995} and reproduce several important features:
-the step change in signal strength from $\omega \leq 6\Omega_i$
-to $\omega > 6\Omega_i$; the
+No. 26148 \cite{Cottrell1993,Dendy1995} and reproduce several important features:
+the step-change in signal strength from $\omega \leq 7\Omega_i$
+to $\omega > 7\Omega_i$; the
 doublet splitting of peaks at $\omega \approx 10\Omega_i$ in Fig.
-\ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\parallel,
-k_\perp)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift
+\ref{fig:2D_F12_physicalpitch}, with origins visible in the fine $(k_\bot,
+k_\parallel)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift
 for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron
@@
 LinearMaxwellVlasov.jl, used to generate the results further enable ICE as
-a diagnostic for fusion plasmas since the code is capable of: calculating the
-dispersion relation of plasmas consisting of an arbitrary number of species; and
+a diagnostic for fusion plasmas since the code is capable of calculating the
+dispersion relation of plasmas consisting of an arbitrary number of species and
 of arbitrary distribution functions, although those with parallel Maxwellians,
-like those presented here, are faster. It has been shown that it is possible to
+like those presented here, are handled faster.
+It has been shown that it is possible to
 qualitatively recreate signals containing lower harmonics of substantial amplitudes by
@@
 PIC simulations has been attributed to nonlinear interactions between pairs of
-more powerful higher harmonics\cite{Carbajal2014}. This work suggests that
+more powerful higher harmonics \cite{Carbajal2014}. This work suggests that
 observations of lower ICE harmonics may arise due to 2D
 linear physics effects but does not exclude the possiblity that nonlinear 1D
-physics\cite{Chapman2018} may also play a part.
+physics \cite{Chapman2018} may also play a part.
 
@@
 plasma giving rise to drifts,
-and the quantisation associatied with toroidal periodicity. These
+and the quantisation arising from toroidal periodicity. These
 effects may create an eigenmode structure, not accounted for herein,
 associated with the kinetic MCI.
-Cottrell et al. \cite{Cottrell1993} does discuss the implications of propagation
-angles on ICE signals and indicate that lower harmonics must propagate
-increasingly close to exactly perpendicular. These effects remain the subject of
-further work to incorporate them in linear and nonlinear studies of ICE.
+Cottrell et al. \cite{Cottrell1993} discusses the implications of propagation
+angles on ICE signals and indicates that lower harmonics must propagate
+increasingly close to exactly perpendicular. It remains the subject of
+further work to incorporate these effects into linear and nonlinear studies of ICE.
 
 This work showcases a new high resolution tool for studying ICE and captures
-some of the fine detail of ICE signals, which is challenging to capture
+some of the fine detail of ICE signals, which are challenging to capture
 with nonlinear PIC
@@
 respectively. As such, this work represents a complementary tool that builds
-upon purely analytical work and nonlinear PIC studies, which in turn enhances
+upon purely analytical work and nonlinear PIC studies to provide
+an explanation of doublet splitting, which in turn enhances
 the feasibility of exploiting ICE as a passive diagnostic of confined and
@@
 
-The plasma dispersion function\cite{Fried1961}
+The plasma dispersion function \cite{Fried1961}
 
@@
 
-\noindent can be calculated via the recurrence relation\cite{Sampoorna2007}
+\noindent can be calculated via the recurrence relation \cite{Sampoorna2007}
 
@@
 Askinazi L G  \textit{et al} 2018 \textit{Nucl. Fusion} \textbf{58} 082003
+
+\bibitem{Fredrickson2019}
+Fredrickson E D \textit{et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
+%Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
+% "Emission in the ion cyclotron range of frequencies (ICE) on NSTX and NSTX-U",
 
@@
 
-\bibitem{Lui2020}
-Liu L \text{et al} 2020 \textit{Nucl. Fusion} \textbf{60} 044002
+\bibitem{Fulop1998}
+F{\"u}l{\"o}p T and Lisak M 1998 \textit{Nucl. Fusion} \textbf{38} 761
 
@@
 % Julia: A fresh approach to numerical computing,
-
-\bibitem{Fredrickson2019}
-Fredrickson E D \textit{et al} 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
-%Fredrickson E D, Gorelenkov  N N, Bell R E, Diallo A, LeBlanc B P, Podestà  M, and NSTX Team 2019 \textit{Physics of Plasmas} \textbf{26}, 032111
-% "Emission in the ion cyclotron range of frequencies (ICE) on NSTX and NSTX-U",
 
@@
 % , A simplex method for function minimization,
-
-\bibitem{Fulop1998}
-F{\"u}l{\"o}p T and Lisak M 1998 \textit{Nucl. Fusion} \textbf{38} 761
 

### 2022-07-28 manuscript.tex
final proof

@@
 \begin{indented}
-\item[]Jan 2022
+\item[]Jul 2022
 \end{indented}
@@
 Linearised fluid approaches that incorporate toroidal geometry and
-inhomogeneous plasma and magnetic fields have shed light of the eigenmode
+inhomogeneous plasma and magnetic fields have shed light on the eigenmode
 structure of compressional Alfv{\'e}n eigenmodes, which as with ICE are also fast
@@
 ring-beam distribution of energetic minority
-alpha-particles, Maxwellian electrons and background thermal majority
+alpha-particles, Maxwellian electrons and Maxwellian majority
 deuteron ions. The ring-beam distribution function, Eq. 1, is a separable
@@
 this distribution function is parameterised by the drift
-velocity parallel to the magnetic field $u_\parallel$, the speed of the
+velocity parallel to the magnetic field $u_\parallel$, the shift of the
 Maxwellian distribution perpendicular to the magnetic
@@
 which $R_0=3.46$ m and
-$\psi_0=55.2^\circ$ are the major radius and pitch angle at birth, and
-$R_{ICE}=4$ m is the major radius of the emission region; see
+$\psi_0=55.2^\circ$ are the major radius and pitch-angle at birth, and
+$R_{ICE}=4$ m is the major radius of the emission region: see
 Figs. 3 and 14 and associated text in Ref. \cite{Cottrell1993}; and
@@
 \begin{align}
-    0 &= ||\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}||,\label{eq:dettensor}\\
+    0 &= |\vec{\vec{\epsilon}} + \left(\vec{k} \otimes \vec{k} - \vec{\vec{I}}(\vec{k}\cdot\vec{k})\right) \frac{c^2}{\omega^2}|,\label{eq:dettensor}\\
     \vec{\vec{\epsilon}} &= \vec{\vec{1}} + \sum_s \frac{\Pi_s^2}{\omega}\sum_{n=-\infty}^{\infty}
@@
 Since all the parallel distribution functions for all species are Maxwellians,
-the parallel integrals can be calculated from the well known plasma dispersion
+the parallel integrals can be calculated from the well-known plasma dispersion
 function\cite{Fried1961}, which is laid out in the appendix.
@@
     real component of the complex $\omega$ solutions), which are annotated along
-    the left hand side and upper edge of the figure. Frequencies are in units of
+    the left-hand side and upper edge of the figure. Frequencies are in units of
     the alpha-particle cyclotron frequency, and wavenumbers are in
@@
 at e.g. the $12^{th}$ harmonic  near $(k_\parallel \simeq 0\; \Omega_{i}/V_A,
-k_\bot \simeq 13\; \Omega_{i}/V_A)$, where $V_A$ is the Alfv{\'e}n speed.
+k_\bot \simeq 13\; \Omega_{i}/V_A)$.
 Qualitatively, scattered points in Fig.
-\ref{fig:2D_F12_physicalpitch}(a) resembles the experimental observations given
+\ref{fig:2D_F12_physicalpitch}(a) resemble the experimental observations given
 by the black trace in that the
-signal below the $7^{th}$ cyclotron harmonics is lower and flatter with well
+signal below the $7^{th}$ cyclotron harmonic is lower and flatter with well
 defined peaks and the signal beyond the $7^{th}$ harmonic is a step-change
-larger and the harmonics are less easily identifiable. These features cannot be
-be recreated with a 1D calculation for these distributions functions
+larger and the harmonics are less easily identifiable. These features cannot
+be recreated with a 1D calculation for these distribution functions
 since they arise from contributions from a range of propagation angles.
@@
 \ref{fig:2D_imag_physicalpitch}, for the case of
-zero pitch angle cosine and is normalised by
+zero pitch-angle cosine and is normalised by
 the high frequency expression for the fast Alfv{\'e}n wave set out in Eq.
@@
 }
-  \right],
-\label{eqn:fastalfven}
+  \right]\label{eqn:fastalfven},
 \end{equation}
@@
     the high frequency expression for the fast Alfv{\'e}n wave set out in
-    Eq. \label{eqn:fastalfven} (Eq. 25 of Ref. \cite{Dendy1994},
+    Eq. \ref{eqn:fastalfven} (Eq. 25 of Ref. \cite{Dendy1994},
     which is also Eq. 11 in Ref. \cite{McClements1996}).
@@
 k_\parallel)$ structure in Fig. \ref{fig:2D_imag_physicalpitch}. The Doppler shift
-for large $k_\parallel$ becomes of the order of the alpha-particle cyclotron
+for large $k_\parallel$ is comparable in magnitude to the alpha-particle cyclotron
 frequency making it difficult to determine which parts of the signal come
-from which cyclotron harmonic numbers, which has important implications for
+from which cyclotron harmonic numbers. This has important implications for
 determining the major radius of origin of ICE.
